### 2013-01-05 paper.tex
initial commit of latex version of paper.  Half of text in.

\documentclass[conference]{IEEEtran}

\usepackage{graphicx,color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{listings}
\usepackage{ifthen}
\usepackage{float} % defines newfloat used below
\usepackage[sort&compress]{natbib} 
%\usepackage{todos} % Use \todoset{all=off} to turn off todos in a paper.
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{microtype}

% hyperref redefines a number of macros, so it should be last.  Empirically,
% doing so eliminates compiler warnings.
\usepackage[bookmarks, colorlinks, citecolor=green, urlcolor=blue,
			filecolor=blue, linkcolor=blue]{hyperref}
% According to the hyperref readme, algorithm must follow hyperref

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Macros

% English
\newcommand{\cf}{\hbox{\emph{cf.}}\xspace}
\newcommand{\deletia}{\ldots [deletia] \ldots}
\newcommand{\etal}{\hbox{\emph{et al.}}\xspace}
\newcommand{\eg}{\hbox{\emph{e.g.}}\xspace}
\newcommand{\ie}{\hbox{\emph{i.e.}}\xspace}
\newcommand{\scil}{\hbox{\emph{sc.}}\xspace} %scilicet: it is permitted to know
\newcommand{\st}{\hbox{\emph{s.t.}}\xspace}
\newcommand{\wrt}{\hbox{\emph{w.r.t.}}\xspace}
\newcommand{\etc}{\hbox{\emph{etc.}}\xspace}
\newcommand{\viz}{\hbox{\emph{viz.}}\xspace} %videlicet: it is permitted to see
\newcommand{\vs}{\hbox{\emph{vs.}}\xspace}

\usepackage{paralist}
\usepackage{mathrsfs}
\usepackage{tabularx}
\usepackage{subfigure}

\newcommand{\ab}[1]{\todo[Alberto]{#1}}
\newcommand{\cb}[1]{\todo[Chris]{#1}}

\renewcommand{\quotation}[1]{``\emph{#1}''}

\iffalse{
\def\todo#1{{\sf $\natural$ #1 $\natural$}}
\def\note#1{{\sf $\clubsuit$ #1 $\clubsuit$}}
\def\marknote#1{{\sf $\spadesuit$ #1 $\spadesuit$}}
}\fi

\usepackage{mdframed}
%% Document-specific hyperref options
\hypersetup{
pdftitle={Expectations, Outcomes, and Challenges of Modern Code Review},
    pdfauthor={Alberto Bacchelli and Christian Bird},
    plainpages=false,
    linkcolor=black, % Overriding these colors to black is somewhat unfortunate 
    %citecolor=black, % b/c the defaults are useful in color.
    citecolor=blue, % b/c the defaults are useful in color.
    filecolor=black,
    urlcolor=black,
    pdfpagelabels
}

%% sig-alternative is stupid
%\pdfpagewidth=8.5in
%\pdfpageheight=11in

% for our nicer-looking \paragraphs under sig-alternate
\newlength{\emstr}
\setlength{\emstr}{0.75em plus 1ex minus 1ex}
\newcommand{\boldpara}[1]{%
\vspace{1.5mm}%
\par\noindent\textbf{#1}\hspace{\emstr}
}%

\newcounter{copyrightbox}

\title{Expectations, Outcomes, and Challenges \\ 
of Modern Code Review}

\author{
	\begin{tabular}{ccc}
		Alberto Bacchelli & \hspace{0.5in} & Christian Bird \\
		REVEAL @ Faculty of Informatics & & Microsoft Research \\
		University of Lugano, Switzerland & & Redmond, Washington, USA \\
		\texttt{alberto.bacchelli@usi.ch} &  &\texttt{cbird@microsoft.com}
	\end{tabular}
}

%\iffalse{
% Only for camera ready -- prevent TeX from rendering widows
%\widowpenalty=9999
%\clubpenalty=9999 
%}\fi

\begin{document}

\maketitle

\begin{abstract}

Code review is a common software engineering practice employed both in open
source and industrial contexts.  Review today is less formal and more
"lightweight" than the code inspections performed and studied in the 70s and
80s. We empirically explore the motivations, challenges, and outcomes of
tool-based code reviews. We observed, interviewed, and surveyed developers and
managers and manually classified hundreds of review comments across diverse
teams at Microsoft. Our study reveals that while finding defects remains the
main motivation for review, reviews are less about defects than expected and
instead provide additional benefits such as knowledge transfer, increased team
awareness, and creation of alternative solutions to problems. Moreover, we find
that code and change understanding is the key aspect of code reviewing and that
developers employ a wide range of mechanisms to meet their understanding needs,
most of which are not met by current tools. We provide recommendations for
practitioners and researchers.

\end{abstract}

%\listoftodos

\section{Introduction}

Peer code review, a manual inspection of source code by developers other than
the author, is recognized as a valuable tool for reducing software defects and
improving the quality of software projects (Ackerman, Fowler and Ebenau 1984)
(Ackerman, Buchwald and Lewski 1989). In 1976, Fagan formalized a highly
structured process for code reviewing (Fagan 1976), based on line-by-line group
reviews, done in extended meetings -- code inspections. Over the years,
researchers provided evidence on code inspection’s benefits, especially in
terms of defect finding, but the cumbersome, time-consuming, and synchronous
nature of this approach hinders its universal adoption in practice (Shull
2008).

Nowadays, many organizations are adopting more lightweight code review
practices to limit the inefficiencies of inspections. In particular, there is a
clear trend toward the usage of tools specifically developed to support code
review (Rigby, Cleary, et al. 2012). In the context of this paper, we define
Modern Code Review, as review that is (1) informal (in contrast to
Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
nowadays, for example at companies such as Microsoft, Google (Kennedy 2006),
Facebook (Tsotsis 2011), and in other companies and OSS projects (Gerrit
(software) 2012).

This trend raises questions, such as: Can we apply the lessons learned from
previous research on code inspections to modern code reviews? What are the
expectations for code review nowadays? What are the actual outcomes of code
review? What challenges do people face in code review?

Answers to these questions can provide insight for both practitioners and
researchers.  Developers and other software project stakeholders can use
empirical evidence about expectations and outcomes to make informed decisions
about when to use code review and how it should fit into their development
process. Researchers can focus their attention on practitioners’ challenges to
make code review more effective.

We present an in-depth study of practices in teams that use modern code review,
revealing what practitioners think, do, and achieve when it comes to modern
code review.

Since Microsoft is made up of many different teams working on very diverse
products, it gives the opportunity to study teams performing code review in
situ and understand their expectations, the benefits they derive from code
review, the needs they have, and the problems they face.

We set up our study as an explorative investigation. We started without a
priori hypotheses regarding how and why code review should be performed, with
the aim of discovering what developers and managers expect from code review,
how reviews are conducted in practice, and what the actual outcomes and
challenges are. To that end, we (1) observed 17 industrial developers
performing code review with various degrees of experience and seniority across
16 separate product teams with distinct reviewing cultures and policies; (2)
interviewed these developers using a semi-structured interviews; (3) manually
inspected and classified the content of 570 comments in discussions contained
within code reviews; and (4) surveyed 165 managers and 873 programmers.

Our results show that, although the top motivation driving code reviews is
still finding defects, the practice and the actual outcomes are less about
finding errors than expected: Defect related comments comprise a small
proportion and mainly cover small logical low-level issues. At the same time,
code review additionally provides a wide spectrum of benefits to software
teams, such as knowledge transfer, team awareness, and improved solutions to
problems. Moreover, we found that context and change understanding is the key
of any review. According to the outcomes they want to achieve, developers
employ many mechanisms to fulfill their understanding needs, most of which are
not currently met by any code review tool.

This paper makes the following contributions:
\begin{itemize}
  \item Characterizing the motivations of developers and managers for code review and compare with actual outcomes.
  \item Relating the outcomes to understanding needs and discuss how developers achieve such needs.
\end{itemize}

Based on our findings, we provide recommendations for practitioners and
implications for researchers as well as outline future avenues for research.

\section{Related Work}

Previous studies exist that have examined the practices of code inspection and
code review.  Stein et al. conducted a study focusing specifically on
distributed, asynchronous code inspections [17]. The study included evaluation
of a tool that allowed for identification and sharing of code faults or
defects. Participants at separated locations can then discuss faults via the
tool. Laitenburger conducted a survey of code inspection methods, and presented
a taxonomy of code inspection techniques (Laitenberger 2002). Johnson conducted
an investigation into code review practices in open source development and the
effect they have on choices made by software project managers (Johnson 2006).

Porter et al. (Porter, Siy and Votta 1996) reported on a review of studies on
code inspection in 1995 that examined the effects of factors such as team size,
type of review, and number of sessions on code inspections.  They also assessed
costs and benefits across a number of studies.  These studies differ from ours
in that they were not tool-based and were the majority involved planned
meetings to discuss the code.

However, prior research also sheds light on why review today is more often
tool-based, informal, and often asynchronous. The current state of code review
might be due to the time required for more formal inspections.  Votta found
that 20\% of the interval in a ``traditional inspection'' is wasted due to
scheduling (Votta 1993). The ICICLE tool (Brothers, Sembugamoorthy and Muller
1990), or ``Intelligent Code Inspection in a C Language Environment,'' was
developed after researchers at Bellcore observed how much time and work was
expended before and during formal code inspections. Many of today’s review
tools are based on ideas that originated in ICICLE.  Other similar tools have
been developed in an effort to reduce time for inspection and allow
asynchronous work on reviews.   Examples include CAIS  (Mashayekhi, Feulner and
Riedl 1994) and Scrutiny (Gintell, et al. 1993).

More recently, Rigby has done extensive work examining code review practices in
open source software development (Rigby, Cleary, et al. 2012).  For example in
a study of practices in the Apache project (Rigby, German and Storey, Open
source software peer review practices: a case study of the apache server 2008)
they data-mined the email archives and found that reviews were typically small
and frequent, and that the contributions to a review were often brief and
independent from one another.

Sutherland and Venolia conducted a study at Microsoft regarding using code
review data for later information needs (Sutherland and Venolia 2009). They
hypothesized that the knowledge exchanged during code reviews could be of great
value to engineers later trying to understand or modify the discussed code.
They found that ``the meat of the code review dialog, no matter what medium, is
the articulation of design rationale'' and, thus, ``code reviews are an enticing
opportunity for capturing design rationale.''

When studying developer work habits, Latoza et al. found that many problems
encountered by developers were related to understanding the rationale behind
code changes and gathering knowledge from other members of their team (LaToza,
Venolia and DeLine 2006).

\section{Methodology}

In this section we define the research questions, describe the research
settings, and outline our research method.

\subsection{Research Questions}

Our investigation of code review revolves around the following research
questions, which we iteratively refined during our initial in-field
observations and interviews:

\begin{enumerate}
  \item What are the motivations and expectations for modern code review? Do they change from managers to developers and testers?
  \item What are the actual outcomes of modern code review? Do they match the expectations?
  \item What are the main challenges experienced when performing modern code reviews relative to the expectations and outcomes?
\end{enumerate}

\subsection{Research Setting}

Our study took place with professional developers, testers, and managers.
Microsoft develops software in diverse domains, from high end server enterprise
data management solutions such as SQL Server to mobile phone applications and
smart phone apps to search engines. Each team has its own development culture
and code review policies. Over the past two years, a common tool for code
review at Microsoft has achieved wide-spread adoption. As it represents a
common and growing solution for code review (over 40,000 developers used it so
far), we focused on developers using this tool for code review -- CodeFlow.

CodeFlow is a collaborative code review tool that allows users to directly
annotate source code in its viewer and interact with review participants in a
live chat model. The functionality of CodeFlow is similar to other review tools
such Google’s Mondrian (Kennedy 2006), Facebook’s Phabricator (Tsotsis 2011) or
open-source  Gerrit (Gerrit (software) 2012). Developers who want their code to
be reviewed create a package with the changed (new, deleted, and modified)
files, select the reviewers, write a message to describe the code review, and
submit everything to the CodeFlow service. CodeFlow then notifies the reviewers
about the incoming task via email.

Once reviewers open a CodeFlow review, they interact with it via a single
desktop window (Figure 1). On the top left (1), they see the list of files
changed in the current submission, plus a ``description.txt'' file, which
contains a textual explanation of the change, written by the author. On bottom
left, CodeFlow shows the list of reviewers and their status (2). We see that
Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
Alberto has reviewed and is waiting for the author to act, as the clock icon
suggests, while Nachi already signed off on the changes. CodeFlow’s main view
(3) shows the diff-highlighted content of the file currently under review. Both
the reviewers and the author can highlight portions of the code and add
comments inline (4). These comments can start threads of discussion and are the
interaction points for the people involved in the review. Each user viewing the
same review in CodeFlow sees events as they happen.  Thus, if an author and
reviewer are working on the review at the same time, the communication is
synchronous and comment threads act similar to instant messaging. The comments
are persisted so that if they work at different times, the communication
becomes asynchronous. The bottom right pane (5) shows the summary of all the
comments in the review.

CodeFlow centralizes and records all the information on code reviews on a
central server. This provides an additional data source that we used to analyze
real code review comments without incurring the Hawthorne effect (Adair 1984).

\subsection{Research Method}

Our research method followed a mixed approach (Creswell 2009), depicted in
Figure 2, collecting data from different sources for triangulation: (1)
analysis of previous study, (2) observations and interviews with developers,
(3) card sort on interview data,  (4) card sort on code review comments, (5)
the creation of an affinity diagram, and (6) survey to managers and
programmers.

1. Analysis of previous study: Our research started with the analysis of a
study commissioned by Microsoft, between April and May 2012 carried out by an
external vendor. The study investigated how different product teams were using
CodeFlow. It consisted of structured interviews (lasting 30-50 minutes) to 23
people with different roles.

Most of the interview questions revolved around topics that are very specific
to tool usage, and were only tangentially related to this work. We found one
relevant as a starting point for our study: ``What do you hope to accomplish
when you submit a code review?'' We analyzed the transcript of this answer, for
each interview, through the process of coding (Berg 2004) (also used in
grounded theory (Adolph 2011)): breaking up the answers into smaller coherent
units (sentences or paragraphs) and adding codes to them. We organized codes
into concepts, which in turn were grouped into more abstract categories.

From this analysis, four motivations emerged for code review: finding defects,
maintaining team awareness, improving code quality, and assessing the
high-level design. We used them to draw an initial guideline for our
interviews.

2. Observations and interviews with developers: Next, we conducted a series of
one-to-one meetings with developers who use CodeFlow, each taking 40-60
minutes.

We contacted 100 random candidates who signed-off between 50 and 250 code
reviews since the CodeFlow release and sampled across different product teams
to address our research questions from a multi-point perspective. We wrote
developers who used CodeFlow in the past and asked them to contact us, giving
us 30 minute notice when they received their next review task so that we could
observe.  The respondents that we interviewed comprised five developers, four
senior developers, six testers, one senior tester, and one software architect.
Their time in the company ranged from 18 months to almost 10 years, with a
median of five years.

Each meeting was comprised of two parts: In the first part, we observed them
performing the code review that they had been assigned. To minimize
invasiveness we used only one observer and to encourage the participant to
narrate their work, we asked the participants to think of us as a newcomer to
the team. In this way, most developers thought aloud without need of prompting.
With consent, we recorded the audio, assuring the participants of anonymity.
Since we, as observers, have backgrounds in software development and practices
at Microsoft, we were able to understand most of the work and where and how
information was obtained without inquiry.

The second part of the meeting was a semi-structured interview (Lindlof and
Taylor 2002). Semi-structured interviews make use of an interview guide that
contains general groupings of topics and questions rather than a pre-determined
exact set and order of questions.  They are often used in an exploratory
context to ``find out what is happening [and] to seek new insights'' (Weiss
1995). The guideline was iteratively refined after each interview, in
particular when developers started providing answers very similar to the
earlier ones, thus reaching a saturation effect.

Observations also reached a saturation point, thus providing insights very similar to the earlier ones. For this, after the first 5-6 observations, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.

The audio of each interview was then transcribed and broken up into smaller
coherent units for subsequent analysis.

3. Card sort (meetings): To group codes that emerged from interviews and
observations into categories, we conducted a card sort. Card sorting is a
sorting technique that is widely used in information architecture to create
mental models and derive taxonomies from input data (Barker 2005). In our case
it helped to organize the codes into hierarchies to deduce a higher level of
abstraction and identify common themes. A card sort involves three phases: In
the (1) preparation phase, participants of the card sort are selected and the
cards are created; in the (2) execution phase, cards are sorted into meaningful
groups with a descriptive title; and in the (3) analysis phase, abstract
hierarchies are formed to deduce general categories.

We applied an open card sort: There were no predefined groups. Instead, the
groups emerged and evolved during the sorting process. In contrast, a closed
card sort has predefined groups and is typically applied when themes are known
in advance, which was not the case for our study.

The first author of this paper created all of the cards, from the 1,047
coherent units in the interviews. Throughout our further analysis other
researchers (the second author and external people) were involved in developing
categories and assigning cards to categories, so as to strengthen the validity
of the result. The first author played a special role of ensuring that the
context of each question was appropriately considered in the categorization,
and creating the initial categories. To ensure the integrity of our categories,
the cards were sorted by the first author several times to identify initial
themes. Next, all researchers reviewed and agreed on the final set of
categories.

4. Card sort (code review comments): The same method was applied to group code
review comments into categories: We randomly sampled 200 threads with at least
two comments (e.g., Point 4 of Figure 2), from the entire dataset of CodeFlow
reviews, which embeds data from dozens of independent software products at
Microsoft. We printed one card for each comment (along with the entire
discussion thread to give the context), totaling 570 cards, and conducted a
card sort, as performed for the interviews, to identify common themes.

5. Affinity Diagram: We used an affinity diagram to organize the categories
that emerged from the card sort. This tool allows large numbers of ideas to be
sorted into groups for review and analysis (Stuart J. Janis 2000). We used it
to generate an overview of the topics that emerged from the card sort, in order
to connect the related concepts and derive the main themes. For generating the
affinity diagram, we followed the five canonical steps: we (1) recorded the
categories on post-it-notes, (2) spread them onto a wall, (3) sorted the
categories based on discussions, until all are sorted and all participants
agreed, (4) named each group with a description, and (5) captured and discussed
the themes. 6. Surveys: The final step of our study was aimed at validating the
concepts that emerged from the previous phases. Towards this goal, we created
two surveys to reach a significant number of participants and to challenge our
conclusions (The full surveys are available as a technical report (Bacchelli
and Bird 2012)). For the design of the surveys, we followed Kitchenham and
Pfleeger’s guidelines for personal opinion surveys (Kitchenham and Pfleeger
2007). Both surveys were anonymous to increase response rates (Tyagi 1989).

We sent the first survey to a cross section of managers.  We considered
managers for which at least half of their team performed code reviews regularly
(on average, one per week or more) and sampled along two dimensions.  The first
dimension was whether or not the manager had participated in a code review
himself since the beginning of the year and the second dimension was whether
the manager managed a single team or multiple teams (a manager of managers).
Thus, we had one sample of first level managers who participated in review,
another sample of second level managers who participated in reviews, etc.  The
first survey was a short survey comprising 6 questions (all optional), which we
sent to 600 managers that had at least ten direct or indirect reporting
developers who used CodeFlow in the past. The central focus was the open
question asking to enumerate the main motivations for doing code reviews in
their team. We received 165 answers (28\% response rate), which we analyzed
before devising the second survey

The second survey comprised 18 questions, mostly closed with multiple choice
answers, and was sent to 2,000 randomly chosen developers who signed off on
average at least one code review per week since the beginning of the year. We
used the time frame of January to June of 2012 to minimize the amount of
organizational churn during the time period and identify employees’ activity in
their current role and team.  We received 873 answers (44\% response rate). Both
response rates were  high, as other online surveys in software engineering have
reported response rates ranging from 14\% to 20\% (Punter, et al. 2003).

\section{Why Do Programmers Do Code Reviews}

Our first research question seeks to understand what motivations and
expectations drive code reviews, and whether managers and developers share the
same opinions.

Based on the responses that we coded from observations of developers performing
code review as well as interviews, there are various motivations for code
review. Overall, the interviews revealed that finding defects, even though
prominent, is just one of the many motivations driving developers to perform
code reviews. Especially when reinforced by a strong team culture around
reviews, developers see code reviews as an activity that has multiple
beneficial influences not only on the code, but also for the team and the
entire development process. In this vein, one senior developer’s comment
summarized many of the responses: \quotation{[code review] also has several
beneficial influences: (1) makes people less protective about their code, (2)
gives another person insight into the code, so there is (3) better sharing of
information across the team, (4) helps support coding conventions on the team,
and [...] (5) helps improving the overall process and quality of code.}

Through the card sort on both meetings and code review comments, we found
several references to motivations for code review and identified six main
topics. To complete this list, in the survey for managers, we included an open
question on why they perform code reviews in their team. We analyzed the
responses to create a comprehensive list of high-level motivations. We included
this list in the developers’ survey and asked them to rank the top three main
reasons that described why they do code reviews.

In the rest of this section, we discuss the motivations that emerged as the
most prominent. We order them according to the importance they were given by
the 873 developers and testers who responded to the final survey.

\subsection{Finding Defects}

One interviewed senior tester explains that he performs code reviews because
they \quotation{are a great source of bugs;} he goes even further stating:
\quotation{sometimes code reviews are a cheaper form of bug finding than
testing.} Moreover, the tool seems not to have an impact on this main
motivation: \quotation{using CodeFlow or using any other tool makes a little
difference to us; it's more about being able to identify flaws in the logic.}

Almost all the managers included ``finding defects'' as one of the reasons for
doing code reviews; for 44\% of the managers, it is the top reason. Managers
considered defects to be both low level issues (e.g., \quotation{correct logic
is in place}) and high level concerns (e.g., \quotation{catch errors in
design}). Concerning surveyed developers/testers, ``finding defects'' is the
first motivation for code review for 383 of the programmers (44\%), second
motivation for 204 (23\%), and third for 96 (11\%).

This is in-line with the reason why code inspections were devised in the first
place: reducing software defects (Ackerman, Fowler and Ebenau 1984).

Nevertheless, even though ``finding defects'' emerged from our data as a strong
motivation (the first for almost half of the programmers and managers),
interviews and survey results indicate that this only tells part of the story
of why practitioners do code reviews and the outcomes they expect.

\subsection{Code Improvement}

Code improvements are comments or changes about code in terms of readability,
commenting, consistency, dead code removal, etc., but do not involve
correctness or defects.

Programmers ranked ``code improvement'' as an important motivation for code
review, close to ``finding defects:'' This is the primary motivation for 337
programmers (39\%), the second for 208 (24\%), and the third for 135 (15\%).
Managers reported code improvements as their primary motivation in 51 cases
(31\%). One manager wrote how code review in her view is a
\quotation{discipline of explaining your code to your peers [that] drives a
higher standard of coding. I think the process is even more important than the
result.}

Most interviewed programmers mentioned that at least one of the reviewers
involved in each code review takes care of checking whether the code follows
the team conventions, for example in terms of code formatting and in terms of
function and variable naming. Some programmers use the ``code improvement''
check as a first step when doing code review: \quotation{the first basic pass
on the code is to check whether it is standard across the team.}

The interviews also gave us a glimpse of the connection between the quality of
code reviews and ``code improvement'' comments. Such comments seem easier to
write and sometimes interviewees mentioned them as the way reviewers use to
avoid spending time to conduct good code reviews. An observation by a senior
developer, in the company for more than nine years, summarizes the opinions we
received from many interviewees: \quotation{I’ve seen quite a few code reviews
where someone commented on formatting while missing the fact that there were
security issues or data model issues.}

\subsection{Alternative Solutions}

``Alternative solutions'' regard changes and comments on improving the
submitted code by adopting an idea that leads to a better implementation. This
is one of the few motivations in which developers and managers do not agree.
While 147 (17\%) developers put this as the first motivation, 202 (23\%) as the
second, and 152 (17\%) as the third, only 4 managers (2\%) even mentioned it
(e.g., \quotation{Generate better ideas, alternative approaches} and
\quotation{Collective wisdom: Someone else on the project may have a better
idea to solve a problem}). The outcome of the interviews was similar to the
position of managers: Interviewees vaguely mentioned this motivation, and
mostly in terms of generic \quotation{better ways to do things.}

\subsection{Knowledge Transfer}

All the interviewees but one motivated their code reviews also from a learning,
or ``knowledge transfer,'' perspective. With the words of a senior developer:
\quotation{one of the things that should be happening with code reviews over
time is a distribution of knowledge. If you do a code review and did not learn
anything about the area and you still do not know anything about the area, then
that was not as good code review as it could have been.} Although we did not
include questions related to ``knowledge transfer'' in our interview guideline,
this topic kept emerging spontaneously from each meeting, thus underscoring its
value for practitioners.

Sometimes programmers told us that they follow code reviews explicitly for
learning purposes. For example, a tester explained: \quotation{[I read code
reviews because] from a code review you can learn about the different parts you
have to touch to implement a certain feature.}

According to interviewees, code review is a learning opportunity for both the
author of the change and the reviewers: There is a bidirectional knowledge
transfer about APIs usage, system design, best practices, team conventions,
\quotation{additional code tricks,} etc. Moreover code reviews are recognized
for educating new developers about code writing.

Managers included ``knowledge transfer'' as one of the reasons for code review,
although never as the top motivation. They mostly wrote about code review as an
education means by mentioning among the motivations for code review:
\quotation{developer education,} \quotation{education for junior developers who
are learning the codebase,} and \quotation{learning tool to teach more junior
team members.}

Programmers answering the survey declared ``knowledge transfer'' to be their
first motivation for code review in 73 cases (8\%), their second in 119 (14\%),
and their third in 141 (16\%).

\subsection{Team Awareness and Transparency}

During one of our observations, one developer was preparing a code review
submission as an author: He wanted other developers to \quotation{double check}
his changes before committing them to the repository. After preparing the code,
he specified the developers he wanted to review his code; he required not only
two specific people, but he also put a generic email distribution group as an
\quotation{optional} reviewer. When we inquired about this choice, he explained
us: \quotation{I am adding [this alias], so that everybody [in the team] is
notified about the change I want to do before I check it in.} In the subsequent
interviews, this concept of using an email list as optional reviewer, or
including specific optional reviewers exclusively for awareness emerged again
frequently, e.g., \quotation{Code reviews are good FYIs [for your
information].}

Managers often mentioned the concept of team awareness as a motivation for code
review, frequently justifying it with the notion of ``transparency:'' Not only
must the team be kept aware of the directions taken by the code, but also
nobody should be allowed to ``secretly'' make changes that might break the code
or alter functionalities.

The 873 programmers answering the survey ranked ``team awareness and
transparency'' very close to ``knowledge transfer.'' In fact, the two concepts
appeared logically related also in the interviews; for example one tester,
while reviewing some code said: \quotation{oh, this guy just implemented this
feature, and now let me back and use it somewhere else.} Showing that he both
learned about the new feature and he was now aware of the possibility to use it
in his own code. 75 (9\%) developers considered team awareness their first
motivation for code review, 108 (12\%) their second, and 149 (17\%) their
third.

Although team awareness and transparency emerged from our data as clearly
promoted by the code review process, academic research seems to have given
little attention to it.

\subsection{Share Code Ownership}

The concept of ``shared code ownership'' is closely related to ``team awareness
and transparency,'' but it has a stronger connotation toward active
collaboration and overlapping coding activities. Programmers and managers
believe that code review is not only an occasion to notify other team members
about incoming changes, but also a means to have more than one knowledgeable
person about specific parts of the codebase. A manager put the following as her
second motivation for code review: \quotation{Broaden knowledge \& understanding
of how specific features/areas are designed and implemented (e.g., grooming
``backup developers'' for areas where knowledge is too concentrated on one or
two expert developers.}

Moreover, both developers and managers have the opinion that practicing code
review also improves the personal perception of team members about shared code
ownership. On this note, a senior developer, with more than 30 years in the
software industry, explained: \quotation{In the past people did not use to do
code reviews and were very reluctant to put themselves in positions where they
were having other people critiquing their code. The fact that code reviews are
considered as a normal thing helps immensely with making people less protective
about their code.} Similarly a manager wrote us explaining that she deems code
reviews important because they \quotation{Dilute any ``rigid sense of
ownership'' that might develop over chunks of code.}

In the programmers’ survey, 51 respondents (6\%) marked ``share code
ownership'' as their first motivation, 100 (11\%) as their second, and (10\%)
as their third.

\subsection{Summary}

In this section, we analyzed the motivations that developers and managers have
for doing code review. We abstracted them into a list, which we finally
included in the programmers’ survey. Figure 3 reports the answers given to this
question: The black bar is the number of developers that put that row as their
top motivation, the gray bar is the number that put it as the second
motivation, etc. We have ordered the factors by giving 3 points for a first
motivation response, 2 points for a second motivation, etc. and then sorting by
the sum.

We discussed the five most prominent motivations, which show that ``finding
defects'' is the top motivation, although participants believe that code review
brings other benefits. The first two motivations were already popular in
research and their effectiveness have been evaluated in the context of code
inspections; on the contrary, the other motivations are still unexplored,
especially those regarding more ``social'' benefits on the team, such as shared
code ownership.

Although motivations are well defined, we still have to verify whether they
actually translate into real outcomes of a modern code review process.

\bibliographystyle{abbrv}
%\setlength{\bibsep}{3pt}
\small
\bibliography{paper}

\end{document}

### 2013-01-05 paper.tex
added codeflow image, and updated paper.tex encoding

@@
 
-\usepackage{graphicx,color}
+\usepackage{graphicx}\graphicspath{{figs/}}
+\usepackage{color}
 \usepackage{amsmath}
@@
 source and industrial contexts.  Review today is less formal and more
-"lightweight" than the code inspections performed and studied in the 70s and
+``lightweight'' than the code inspections performed and studied in the 70s and
 80s. We empirically explore the motivations, challenges, and outcomes of
@@
 reviews, done in extended meetings -- code inspections. Over the years,
-researchers provided evidence on code inspection’s benefits, especially in
+researchers provided evidence on code inspectionís benefits, especially in
 terms of defect finding, but the cumbersome, time-consuming, and synchronous
@@
 about when to use code review and how it should fit into their development
-process. Researchers can focus their attention on practitioners’ challenges to
+process. Researchers can focus their attention on practitionersí challenges to
 make code review more effective.
@@
 developed after researchers at Bellcore observed how much time and work was
-expended before and during formal code inspections. Many of today’s review
+expended before and during formal code inspections. Many of todayís review
 tools are based on ideas that originated in ICICLE.  Other similar tools have
@@
 live chat model. The functionality of CodeFlow is similar to other review tools
-such Google’s Mondrian (Kennedy 2006), Facebook’s Phabricator (Tsotsis 2011) or
+such Googleís Mondrian (Kennedy 2006), Facebookís Phabricator (Tsotsis 2011) or
 open-source  Gerrit (Gerrit (software) 2012). Developers who want their code to
@@
 Alberto has reviewed and is waiting for the author to act, as the clock icon
-suggests, while Nachi already signed off on the changes. CodeFlow’s main view
+suggests, while Nachi already signed off on the changes. CodeFlowís main view
 (3) shows the diff-highlighted content of the file currently under review. Both
@@
 and Bird 2012)). For the design of the surveys, we followed Kitchenham and
-Pfleeger’s guidelines for personal opinion surveys (Kitchenham and Pfleeger
+Pfleegerís guidelines for personal opinion surveys (Kitchenham and Pfleeger
 2007). Both surveys were anonymous to increase response rates (Tyagi 1989).
@@
 used the time frame of January to June of 2012 to minimize the amount of
-organizational churn during the time period and identify employees’ activity in
+organizational churn during the time period and identify employeesí activity in
 their current role and team.  We received 873 answers (44\% response rate). Both
@@
 beneficial influences not only on the code, but also for the team and the
-entire development process. In this vein, one senior developer’s comment
+entire development process. In this vein, one senior developerís comment
 summarized many of the responses: \quotation{[code review] also has several
@@
 responses to create a comprehensive list of high-level motivations. We included
-this list in the developers’ survey and asked them to rank the top three main
+this list in the developersí survey and asked them to rank the top three main
 reasons that described why they do code reviews.
@@
 developer, in the company for more than nine years, summarizes the opinions we
-received from many interviewees: \quotation{I’ve seen quite a few code reviews
+received from many interviewees: \quotation{Iíve seen quite a few code reviews
 where someone commented on formatting while missing the fact that there were
@@
 
-In the programmers’ survey, 51 respondents (6\%) marked ``share code
+In the programmersí survey, 51 respondents (6\%) marked ``share code
 ownership'' as their first motivation, 100 (11\%) as their second, and (10\%)
@@
 for doing code review. We abstracted them into a list, which we finally
-included in the programmers’ survey. Figure 3 reports the answers given to this
+included in the programmersí survey. Figure 3 reports the answers given to this
 question: The black bar is the number of developers that put that row as their

### 2013-01-05 paper.tex
split the document in one file per section, so that it is easier to manage in more than one author

@@
 
-\section{Introduction}
-
-Peer code review, a manual inspection of source code by developers other than
-the author, is recognized as a valuable tool for reducing software defects and
-improving the quality of software projects (Ackerman, Fowler and Ebenau 1984)
-(Ackerman, Buchwald and Lewski 1989). In 1976, Fagan formalized a highly
-structured process for code reviewing (Fagan 1976), based on line-by-line group
-reviews, done in extended meetings -- code inspections. Over the years,
-researchers provided evidence on code inspectionís benefits, especially in
-terms of defect finding, but the cumbersome, time-consuming, and synchronous
-nature of this approach hinders its universal adoption in practice (Shull
-2008).
-
-Nowadays, many organizations are adopting more lightweight code review
-practices to limit the inefficiencies of inspections. In particular, there is a
-clear trend toward the usage of tools specifically developed to support code
-review (Rigby, Cleary, et al. 2012). In the context of this paper, we define
-Modern Code Review, as review that is (1) informal (in contrast to
-Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
-nowadays, for example at companies such as Microsoft, Google (Kennedy 2006),
-Facebook (Tsotsis 2011), and in other companies and OSS projects (Gerrit
-(software) 2012).
-
-This trend raises questions, such as: Can we apply the lessons learned from
-previous research on code inspections to modern code reviews? What are the
-expectations for code review nowadays? What are the actual outcomes of code
-review? What challenges do people face in code review?
-
-Answers to these questions can provide insight for both practitioners and
-researchers.  Developers and other software project stakeholders can use
-empirical evidence about expectations and outcomes to make informed decisions
-about when to use code review and how it should fit into their development
-process. Researchers can focus their attention on practitionersí challenges to
-make code review more effective.
-
-We present an in-depth study of practices in teams that use modern code review,
-revealing what practitioners think, do, and achieve when it comes to modern
-code review.
-
-Since Microsoft is made up of many different teams working on very diverse
-products, it gives the opportunity to study teams performing code review in
-situ and understand their expectations, the benefits they derive from code
-review, the needs they have, and the problems they face.
-
-We set up our study as an explorative investigation. We started without a
-priori hypotheses regarding how and why code review should be performed, with
-the aim of discovering what developers and managers expect from code review,
-how reviews are conducted in practice, and what the actual outcomes and
-challenges are. To that end, we (1) observed 17 industrial developers
-performing code review with various degrees of experience and seniority across
-16 separate product teams with distinct reviewing cultures and policies; (2)
-interviewed these developers using a semi-structured interviews; (3) manually
-inspected and classified the content of 570 comments in discussions contained
-within code reviews; and (4) surveyed 165 managers and 873 programmers.
-
-Our results show that, although the top motivation driving code reviews is
-still finding defects, the practice and the actual outcomes are less about
-finding errors than expected: Defect related comments comprise a small
-proportion and mainly cover small logical low-level issues. At the same time,
-code review additionally provides a wide spectrum of benefits to software
-teams, such as knowledge transfer, team awareness, and improved solutions to
-problems. Moreover, we found that context and change understanding is the key
-of any review. According to the outcomes they want to achieve, developers
-employ many mechanisms to fulfill their understanding needs, most of which are
-not currently met by any code review tool.
-
-This paper makes the following contributions:
-\begin{itemize}
-  \item Characterizing the motivations of developers and managers for code review and compare with actual outcomes.
-  \item Relating the outcomes to understanding needs and discuss how developers achieve such needs.
-\end{itemize}
-
-Based on our findings, we provide recommendations for practitioners and
-implications for researchers as well as outline future avenues for research.
-
-\section{Related Work}
-
-Previous studies exist that have examined the practices of code inspection and
-code review.  Stein et al. conducted a study focusing specifically on
-distributed, asynchronous code inspections [17]. The study included evaluation
-of a tool that allowed for identification and sharing of code faults or
-defects. Participants at separated locations can then discuss faults via the
-tool. Laitenburger conducted a survey of code inspection methods, and presented
-a taxonomy of code inspection techniques (Laitenberger 2002). Johnson conducted
-an investigation into code review practices in open source development and the
-effect they have on choices made by software project managers (Johnson 2006).
-
-Porter et al. (Porter, Siy and Votta 1996) reported on a review of studies on
-code inspection in 1995 that examined the effects of factors such as team size,
-type of review, and number of sessions on code inspections.  They also assessed
-costs and benefits across a number of studies.  These studies differ from ours
-in that they were not tool-based and were the majority involved planned
-meetings to discuss the code.
-
-However, prior research also sheds light on why review today is more often
-tool-based, informal, and often asynchronous. The current state of code review
-might be due to the time required for more formal inspections.  Votta found
-that 20\% of the interval in a ``traditional inspection'' is wasted due to
-scheduling (Votta 1993). The ICICLE tool (Brothers, Sembugamoorthy and Muller
-1990), or ``Intelligent Code Inspection in a C Language Environment,'' was
-developed after researchers at Bellcore observed how much time and work was
-expended before and during formal code inspections. Many of todayís review
-tools are based on ideas that originated in ICICLE.  Other similar tools have
-been developed in an effort to reduce time for inspection and allow
-asynchronous work on reviews.   Examples include CAIS  (Mashayekhi, Feulner and
-Riedl 1994) and Scrutiny (Gintell, et al. 1993).
-
-More recently, Rigby has done extensive work examining code review practices in
-open source software development (Rigby, Cleary, et al. 2012).  For example in
-a study of practices in the Apache project (Rigby, German and Storey, Open
-source software peer review practices: a case study of the apache server 2008)
-they data-mined the email archives and found that reviews were typically small
-and frequent, and that the contributions to a review were often brief and
-independent from one another.
-
-Sutherland and Venolia conducted a study at Microsoft regarding using code
-review data for later information needs (Sutherland and Venolia 2009). They
-hypothesized that the knowledge exchanged during code reviews could be of great
-value to engineers later trying to understand or modify the discussed code.
-They found that ``the meat of the code review dialog, no matter what medium, is
-the articulation of design rationale'' and, thus, ``code reviews are an enticing
-opportunity for capturing design rationale.''
-
-When studying developer work habits, Latoza et al. found that many problems
-encountered by developers were related to understanding the rationale behind
-code changes and gathering knowledge from other members of their team (LaToza,
-Venolia and DeLine 2006).
-
-\section{Methodology}
-
-In this section we define the research questions, describe the research
-settings, and outline our research method.
-
-\subsection{Research Questions}
-
-Our investigation of code review revolves around the following research
-questions, which we iteratively refined during our initial in-field
-observations and interviews:
-
-\begin{enumerate}
-  \item What are the motivations and expectations for modern code review? Do they change from managers to developers and testers?
-  \item What are the actual outcomes of modern code review? Do they match the expectations?
-  \item What are the main challenges experienced when performing modern code reviews relative to the expectations and outcomes?
-\end{enumerate}
-
-\subsection{Research Setting}
-
-Our study took place with professional developers, testers, and managers.
-Microsoft develops software in diverse domains, from high end server enterprise
-data management solutions such as SQL Server to mobile phone applications and
-smart phone apps to search engines. Each team has its own development culture
-and code review policies. Over the past two years, a common tool for code
-review at Microsoft has achieved wide-spread adoption. As it represents a
-common and growing solution for code review (over 40,000 developers used it so
-far), we focused on developers using this tool for code review -- CodeFlow.
-
-CodeFlow is a collaborative code review tool that allows users to directly
-annotate source code in its viewer and interact with review participants in a
-live chat model. The functionality of CodeFlow is similar to other review tools
-such Googleís Mondrian (Kennedy 2006), Facebookís Phabricator (Tsotsis 2011) or
-open-source  Gerrit (Gerrit (software) 2012). Developers who want their code to
-be reviewed create a package with the changed (new, deleted, and modified)
-files, select the reviewers, write a message to describe the code review, and
-submit everything to the CodeFlow service. CodeFlow then notifies the reviewers
-about the incoming task via email.
-
-Once reviewers open a CodeFlow review, they interact with it via a single
-desktop window (Figure 1). On the top left (1), they see the list of files
-changed in the current submission, plus a ``description.txt'' file, which
-contains a textual explanation of the change, written by the author. On bottom
-left, CodeFlow shows the list of reviewers and their status (2). We see that
-Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
-Alberto has reviewed and is waiting for the author to act, as the clock icon
-suggests, while Nachi already signed off on the changes. CodeFlowís main view
-(3) shows the diff-highlighted content of the file currently under review. Both
-the reviewers and the author can highlight portions of the code and add
-comments inline (4). These comments can start threads of discussion and are the
-interaction points for the people involved in the review. Each user viewing the
-same review in CodeFlow sees events as they happen.  Thus, if an author and
-reviewer are working on the review at the same time, the communication is
-synchronous and comment threads act similar to instant messaging. The comments
-are persisted so that if they work at different times, the communication
-becomes asynchronous. The bottom right pane (5) shows the summary of all the
-comments in the review.
-
-CodeFlow centralizes and records all the information on code reviews on a
-central server. This provides an additional data source that we used to analyze
-real code review comments without incurring the Hawthorne effect (Adair 1984).
-
-\subsection{Research Method}
-
-Our research method followed a mixed approach (Creswell 2009), depicted in
-Figure 2, collecting data from different sources for triangulation: (1)
-analysis of previous study, (2) observations and interviews with developers,
-(3) card sort on interview data,  (4) card sort on code review comments, (5)
-the creation of an affinity diagram, and (6) survey to managers and
-programmers.
-
-1. Analysis of previous study: Our research started with the analysis of a
-study commissioned by Microsoft, between April and May 2012 carried out by an
-external vendor. The study investigated how different product teams were using
-CodeFlow. It consisted of structured interviews (lasting 30-50 minutes) to 23
-people with different roles.
-
-Most of the interview questions revolved around topics that are very specific
-to tool usage, and were only tangentially related to this work. We found one
-relevant as a starting point for our study: ``What do you hope to accomplish
-when you submit a code review?'' We analyzed the transcript of this answer, for
-each interview, through the process of coding (Berg 2004) (also used in
-grounded theory (Adolph 2011)): breaking up the answers into smaller coherent
-units (sentences or paragraphs) and adding codes to them. We organized codes
-into concepts, which in turn were grouped into more abstract categories.
-
-From this analysis, four motivations emerged for code review: finding defects,
-maintaining team awareness, improving code quality, and assessing the
-high-level design. We used them to draw an initial guideline for our
-interviews.
-
-2. Observations and interviews with developers: Next, we conducted a series of
-one-to-one meetings with developers who use CodeFlow, each taking 40-60
-minutes.
-
-We contacted 100 random candidates who signed-off between 50 and 250 code
-reviews since the CodeFlow release and sampled across different product teams
-to address our research questions from a multi-point perspective. We wrote
-developers who used CodeFlow in the past and asked them to contact us, giving
-us 30 minute notice when they received their next review task so that we could
-observe.  The respondents that we interviewed comprised five developers, four
-senior developers, six testers, one senior tester, and one software architect.
-Their time in the company ranged from 18 months to almost 10 years, with a
-median of five years.
-
-Each meeting was comprised of two parts: In the first part, we observed them
-performing the code review that they had been assigned. To minimize
-invasiveness we used only one observer and to encourage the participant to
-narrate their work, we asked the participants to think of us as a newcomer to
-the team. In this way, most developers thought aloud without need of prompting.
-With consent, we recorded the audio, assuring the participants of anonymity.
-Since we, as observers, have backgrounds in software development and practices
-at Microsoft, we were able to understand most of the work and where and how
-information was obtained without inquiry.
-
-The second part of the meeting was a semi-structured interview (Lindlof and
-Taylor 2002). Semi-structured interviews make use of an interview guide that
-contains general groupings of topics and questions rather than a pre-determined
-exact set and order of questions.  They are often used in an exploratory
-context to ``find out what is happening [and] to seek new insights'' (Weiss
-1995). The guideline was iteratively refined after each interview, in
-particular when developers started providing answers very similar to the
-earlier ones, thus reaching a saturation effect.
-
-Observations also reached a saturation point, thus providing insights very similar to the earlier ones.
-For this, after the first 5-6 observations, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.
-
-The audio of each interview was then transcribed and broken up into smaller
-coherent units for subsequent analysis.
-
-3. Card sort (meetings): To group codes that emerged from interviews and
-observations into categories, we conducted a card sort. Card sorting is a
-sorting technique that is widely used in information architecture to create
-mental models and derive taxonomies from input data (Barker 2005). In our case
-it helped to organize the codes into hierarchies to deduce a higher level of
-abstraction and identify common themes. A card sort involves three phases: In
-the (1) preparation phase, participants of the card sort are selected and the
-cards are created; in the (2) execution phase, cards are sorted into meaningful
-groups with a descriptive title; and in the (3) analysis phase, abstract
-hierarchies are formed to deduce general categories.
-
-We applied an open card sort: There were no predefined groups. Instead, the
-groups emerged and evolved during the sorting process. In contrast, a closed
-card sort has predefined groups and is typically applied when themes are known
-in advance, which was not the case for our study.
-
-The first author of this paper created all of the cards, from the 1,047
-coherent units in the interviews. Throughout our further analysis other
-researchers (the second author and external people) were involved in developing
-categories and assigning cards to categories, so as to strengthen the validity
-of the result. The first author played a special role of ensuring that the
-context of each question was appropriately considered in the categorization,
-and creating the initial categories. To ensure the integrity of our categories,
-the cards were sorted by the first author several times to identify initial
-themes. Next, all researchers reviewed and agreed on the final set of
-categories.
-
-4. Card sort (code review comments): The same method was applied to group code
-review comments into categories: We randomly sampled 200 threads with at least
-two comments (e.g., Point 4 of Figure 2), from the entire dataset of CodeFlow
-reviews, which embeds data from dozens of independent software products at
-Microsoft. We printed one card for each comment (along with the entire
-discussion thread to give the context), totaling 570 cards, and conducted a
-card sort, as performed for the interviews, to identify common themes.
-
-5. Affinity Diagram: We used an affinity diagram to organize the categories
-that emerged from the card sort. This tool allows large numbers of ideas to be
-sorted into groups for review and analysis (Stuart J. Janis 2000). We used it
-to generate an overview of the topics that emerged from the card sort, in order
-to connect the related concepts and derive the main themes. For generating the
-affinity diagram, we followed the five canonical steps: we (1) recorded the
-categories on post-it-notes, (2) spread them onto a wall, (3) sorted the
-categories based on discussions, until all are sorted and all participants
-agreed, (4) named each group with a description, and (5) captured and discussed
-the themes. 6. Surveys: The final step of our study was aimed at validating the
-concepts that emerged from the previous phases. Towards this goal, we created
-two surveys to reach a significant number of participants and to challenge our
-conclusions (The full surveys are available as a technical report (Bacchelli
-and Bird 2012)). For the design of the surveys, we followed Kitchenham and
-Pfleegerís guidelines for personal opinion surveys (Kitchenham and Pfleeger
-2007). Both surveys were anonymous to increase response rates (Tyagi 1989).
-
-We sent the first survey to a cross section of managers.  We considered
-managers for which at least half of their team performed code reviews regularly
-(on average, one per week or more) and sampled along two dimensions.  The first
-dimension was whether or not the manager had participated in a code review
-himself since the beginning of the year and the second dimension was whether
-the manager managed a single team or multiple teams (a manager of managers).
-Thus, we had one sample of first level managers who participated in review,
-another sample of second level managers who participated in reviews, etc.  The
-first survey was a short survey comprising 6 questions (all optional), which we
-sent to 600 managers that had at least ten direct or indirect reporting
-developers who used CodeFlow in the past. The central focus was the open
-question asking to enumerate the main motivations for doing code reviews in
-their team. We received 165 answers (28\% response rate), which we analyzed
-before devising the second survey
-
-The second survey comprised 18 questions, mostly closed with multiple choice
-answers, and was sent to 2,000 randomly chosen developers who signed off on
-average at least one code review per week since the beginning of the year. We
-used the time frame of January to June of 2012 to minimize the amount of
-organizational churn during the time period and identify employeesí activity in
-their current role and team.  We received 873 answers (44\% response rate). Both
-response rates were  high, as other online surveys in software engineering have
-reported response rates ranging from 14\% to 20\% (Punter, et al. 2003).
-
-\section{Why Do Programmers Do Code Reviews}
-
-Our first research question seeks to understand what motivations and
-expectations drive code reviews, and whether managers and developers share the
-same opinions.
-
-Based on the responses that we coded from observations of developers performing
-code review as well as interviews, there are various motivations for code
-review. Overall, the interviews revealed that finding defects, even though
-prominent, is just one of the many motivations driving developers to perform
-code reviews. Especially when reinforced by a strong team culture around
-reviews, developers see code reviews as an activity that has multiple
-beneficial influences not only on the code, but also for the team and the
-entire development process. In this vein, one senior developerís comment
-summarized many of the responses: \quotation{[code review] also has several
-beneficial influences: (1) makes people less protective about their code, (2)
-gives another person insight into the code, so there is (3) better sharing of
-information across the team, (4) helps support coding conventions on the team,
-and [...] (5) helps improving the overall process and quality of code.}
-
-Through the card sort on both meetings and code review comments, we found
-several references to motivations for code review and identified six main
-topics. To complete this list, in the survey for managers, we included an open
-question on why they perform code reviews in their team. We analyzed the
-responses to create a comprehensive list of high-level motivations. We included
-this list in the developersí survey and asked them to rank the top three main
-reasons that described why they do code reviews.
-
-In the rest of this section, we discuss the motivations that emerged as the
-most prominent. We order them according to the importance they were given by
-the 873 developers and testers who responded to the final survey.
-
-\subsection{Finding Defects}
-
-One interviewed senior tester explains that he performs code reviews because
-they \quotation{are a great source of bugs;} he goes even further stating:
-\quotation{sometimes code reviews are a cheaper form of bug finding than
-testing.} Moreover, the tool seems not to have an impact on this main
-motivation: \quotation{using CodeFlow or using any other tool makes a little
-difference to us; it's more about being able to identify flaws in the logic.}
-
-Almost all the managers included ``finding defects'' as one of the reasons for
-doing code reviews; for 44\% of the managers, it is the top reason. Managers
-considered defects to be both low level issues (e.g., \quotation{correct logic
-is in place}) and high level concerns (e.g., \quotation{catch errors in
-design}). Concerning surveyed developers/testers, ``finding defects'' is the
-first motivation for code review for 383 of the programmers (44\%), second
-motivation for 204 (23\%), and third for 96 (11\%).
-
-This is in-line with the reason why code inspections were devised in the first
-place: reducing software defects (Ackerman, Fowler and Ebenau 1984).
-
-Nevertheless, even though ``finding defects'' emerged from our data as a strong
-motivation (the first for almost half of the programmers and managers),
-interviews and survey results indicate that this only tells part of the story
-of why practitioners do code reviews and the outcomes they expect.
-
-\subsection{Code Improvement}
-
-Code improvements are comments or changes about code in terms of readability,
-commenting, consistency, dead code removal, etc., but do not involve
-correctness or defects.
-
-Programmers ranked ``code improvement'' as an important motivation for code
-review, close to ``finding defects:'' This is the primary motivation for 337
-programmers (39\%), the second for 208 (24\%), and the third for 135 (15\%).
-Managers reported code improvements as their primary motivation in 51 cases
-(31\%). One manager wrote how code review in her view is a
-\quotation{discipline of explaining your code to your peers [that] drives a
-higher standard of coding. I think the process is even more important than the
-result.}
-
-Most interviewed programmers mentioned that at least one of the reviewers
-involved in each code review takes care of checking whether the code follows
-the team conventions, for example in terms of code formatting and in terms of
-function and variable naming. Some programmers use the ``code improvement''
-check as a first step when doing code review: \quotation{the first basic pass
-on the code is to check whether it is standard across the team.}
-
-The interviews also gave us a glimpse of the connection between the quality of
-code reviews and ``code improvement'' comments. Such comments seem easier to
-write and sometimes interviewees mentioned them as the way reviewers use to
-avoid spending time to conduct good code reviews. An observation by a senior
-developer, in the company for more than nine years, summarizes the opinions we
-received from many interviewees: \quotation{Iíve seen quite a few code reviews
-where someone commented on formatting while missing the fact that there were
-security issues or data model issues.}
-
-\subsection{Alternative Solutions}
-
-``Alternative solutions'' regard changes and comments on improving the
-submitted code by adopting an idea that leads to a better implementation. This
-is one of the few motivations in which developers and managers do not agree.
-While 147 (17\%) developers put this as the first motivation, 202 (23\%) as the
-second, and 152 (17\%) as the third, only 4 managers (2\%) even mentioned it
-(e.g., \quotation{Generate better ideas, alternative approaches} and
-\quotation{Collective wisdom: Someone else on the project may have a better
-idea to solve a problem}). The outcome of the interviews was similar to the
-position of managers: Interviewees vaguely mentioned this motivation, and
-mostly in terms of generic \quotation{better ways to do things.}
-
-\subsection{Knowledge Transfer}
-
-All the interviewees but one motivated their code reviews also from a learning,
-or ``knowledge transfer,'' perspective. With the words of a senior developer:
-\quotation{one of the things that should be happening with code reviews over
-time is a distribution of knowledge. If you do a code review and did not learn
-anything about the area and you still do not know anything about the area, then
-that was not as good code review as it could have been.} Although we did not
-include questions related to ``knowledge transfer'' in our interview guideline,
-this topic kept emerging spontaneously from each meeting, thus underscoring its
-value for practitioners.
-
-Sometimes programmers told us that they follow code reviews explicitly for
-learning purposes. For example, a tester explained: \quotation{[I read code
-reviews because] from a code review you can learn about the different parts you
-have to touch to implement a certain feature.}
-
-According to interviewees, code review is a learning opportunity for both the
-author of the change and the reviewers: There is a bidirectional knowledge
-transfer about APIs usage, system design, best practices, team conventions,
-\quotation{additional code tricks,} etc. Moreover code reviews are recognized
-for educating new developers about code writing.
-
-Managers included ``knowledge transfer'' as one of the reasons for code review,
-although never as the top motivation. They mostly wrote about code review as an
-education means by mentioning among the motivations for code review:
-\quotation{developer education,} \quotation{education for junior developers who
-are learning the codebase,} and \quotation{learning tool to teach more junior
-team members.}
-
-Programmers answering the survey declared ``knowledge transfer'' to be their
-first motivation for code review in 73 cases (8\%), their second in 119 (14\%),
-and their third in 141 (16\%).
-
-\subsection{Team Awareness and Transparency}
-
-During one of our observations, one developer was preparing a code review
-submission as an author: He wanted other developers to \quotation{double check}
-his changes before committing them to the repository. After preparing the code,
-he specified the developers he wanted to review his code; he required not only
-two specific people, but he also put a generic email distribution group as an
-\quotation{optional} reviewer. When we inquired about this choice, he explained
-us: \quotation{I am adding [this alias], so that everybody [in the team] is
-notified about the change I want to do before I check it in.} In the subsequent
-interviews, this concept of using an email list as optional reviewer, or
-including specific optional reviewers exclusively for awareness emerged again
-frequently, e.g., \quotation{Code reviews are good FYIs [for your
-information].}
-
-Managers often mentioned the concept of team awareness as a motivation for code
-review, frequently justifying it with the notion of ``transparency:'' Not only
-must the team be kept aware of the directions taken by the code, but also
-nobody should be allowed to ``secretly'' make changes that might break the code
-or alter functionalities.
-
-The 873 programmers answering the survey ranked ``team awareness and
-transparency'' very close to ``knowledge transfer.'' In fact, the two concepts
-appeared logically related also in the interviews; for example one tester,
-while reviewing some code said: \quotation{oh, this guy just implemented this
-feature, and now let me back and use it somewhere else.} Showing that he both
-learned about the new feature and he was now aware of the possibility to use it
-in his own code. 75 (9\%) developers considered team awareness their first
-motivation for code review, 108 (12\%) their second, and 149 (17\%) their
-third.
-
-Although team awareness and transparency emerged from our data as clearly
-promoted by the code review process, academic research seems to have given
-little attention to it.
-
-\subsection{Share Code Ownership}
-
-The concept of ``shared code ownership'' is closely related to ``team awareness
-and transparency,'' but it has a stronger connotation toward active
-collaboration and overlapping coding activities. Programmers and managers
-believe that code review is not only an occasion to notify other team members
-about incoming changes, but also a means to have more than one knowledgeable
-person about specific parts of the codebase. A manager put the following as her
-second motivation for code review: \quotation{Broaden knowledge \& understanding
-of how specific features/areas are designed and implemented (e.g., grooming
-``backup developers'' for areas where knowledge is too concentrated on one or
-two expert developers.}
-
-Moreover, both developers and managers have the opinion that practicing code
-review also improves the personal perception of team members about shared code
-ownership. On this note, a senior developer, with more than 30 years in the
-software industry, explained: \quotation{In the past people did not use to do
-code reviews and were very reluctant to put themselves in positions where they
-were having other people critiquing their code. The fact that code reviews are
-considered as a normal thing helps immensely with making people less protective
-about their code.} Similarly a manager wrote us explaining that she deems code
-reviews important because they \quotation{Dilute any ``rigid sense of
-ownership'' that might develop over chunks of code.}
-
-In the programmersí survey, 51 respondents (6\%) marked ``share code
-ownership'' as their first motivation, 100 (11\%) as their second, and (10\%)
-as their third.
-
-\subsection{Summary}
-
-In this section, we analyzed the motivations that developers and managers have
-for doing code review. We abstracted them into a list, which we finally
-included in the programmersí survey. Figure 3 reports the answers given to this
-question: The black bar is the number of developers that put that row as their
-top motivation, the gray bar is the number that put it as the second
-motivation, etc. We have ordered the factors by giving 3 points for a first
-motivation response, 2 points for a second motivation, etc. and then sorting by
-the sum.
-
-We discussed the five most prominent motivations, which show that ``finding
-defects'' is the top motivation, although participants believe that code review
-brings other benefits. The first two motivations were already popular in
-research and their effectiveness have been evaluated in the context of code
-inspections; on the contrary, the other motivations are still unexplored,
-especially those regarding more ``social'' benefits on the team, such as shared
-code ownership.
-
-Although motivations are well defined, we still have to verify whether they
-actually translate into real outcomes of a modern code review process.
+\input{introduction}
+\input{related_work}
+\input{methodology}
+\input{expectations}
+\input{outcomes}
+\input{challenges}
+\input{implications}
+\input{limitations}
+\input{conclusion}
 

### 2013-01-05 sections/challenges.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%

### 2013-01-05 sections/expectations.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Why Do Programmers Do Code Reviews} \label{sec:expectations}
%%%%%%%%%%%%%%%%%%%%%%

Our first research question seeks to understand what motivations and
expectations drive code reviews, and whether managers and developers share the
same opinions.

Based on the responses that we coded from observations of developers performing
code review as well as interviews, there are various motivations for code
review. Overall, the interviews revealed that finding defects, even though
prominent, is just one of the many motivations driving developers to perform
code reviews. Especially when reinforced by a strong team culture around
reviews, developers see code reviews as an activity that has multiple
beneficial influences not only on the code, but also for the team and the
entire development process. In this vein, one senior developerís comment
summarized many of the responses: \quotation{[code review] also has several
beneficial influences: (1) makes people less protective about their code, (2)
gives another person insight into the code, so there is (3) better sharing of
information across the team, (4) helps support coding conventions on the team,
and [...] (5) helps improving the overall process and quality of code.}

Through the card sort on both meetings and code review comments, we found
several references to motivations for code review and identified six main
topics. To complete this list, in the survey for managers, we included an open
question on why they perform code reviews in their team. We analyzed the
responses to create a comprehensive list of high-level motivations. We included
this list in the developersí survey and asked them to rank the top three main
reasons that described why they do code reviews.

In the rest of this section, we discuss the motivations that emerged as the
most prominent. We order them according to the importance they were given by
the 873 developers and testers who responded to the final survey.

\subsection{Finding Defects}

One interviewed senior tester explains that he performs code reviews because
they \quotation{are a great source of bugs;} he goes even further stating:
\quotation{sometimes code reviews are a cheaper form of bug finding than
testing.} Moreover, the tool seems not to have an impact on this main
motivation: \quotation{using CodeFlow or using any other tool makes a little
difference to us; it's more about being able to identify flaws in the logic.}

Almost all the managers included ``finding defects'' as one of the reasons for
doing code reviews; for 44\% of the managers, it is the top reason. Managers
considered defects to be both low level issues (e.g., \quotation{correct logic
is in place}) and high level concerns (e.g., \quotation{catch errors in
design}). Concerning surveyed developers/testers, ``finding defects'' is the
first motivation for code review for 383 of the programmers (44\%), second
motivation for 204 (23\%), and third for 96 (11\%).

This is in-line with the reason why code inspections were devised in the first
place: reducing software defects (Ackerman, Fowler and Ebenau 1984).

Nevertheless, even though ``finding defects'' emerged from our data as a strong
motivation (the first for almost half of the programmers and managers),
interviews and survey results indicate that this only tells part of the story
of why practitioners do code reviews and the outcomes they expect.

\subsection{Code Improvement}

Code improvements are comments or changes about code in terms of readability,
commenting, consistency, dead code removal, etc., but do not involve
correctness or defects.

Programmers ranked ``code improvement'' as an important motivation for code
review, close to ``finding defects:'' This is the primary motivation for 337
programmers (39\%), the second for 208 (24\%), and the third for 135 (15\%).
Managers reported code improvements as their primary motivation in 51 cases
(31\%). One manager wrote how code review in her view is a
\quotation{discipline of explaining your code to your peers [that] drives a
higher standard of coding. I think the process is even more important than the
result.}

Most interviewed programmers mentioned that at least one of the reviewers
involved in each code review takes care of checking whether the code follows
the team conventions, for example in terms of code formatting and in terms of
function and variable naming. Some programmers use the ``code improvement''
check as a first step when doing code review: \quotation{the first basic pass
on the code is to check whether it is standard across the team.}

The interviews also gave us a glimpse of the connection between the quality of
code reviews and ``code improvement'' comments. Such comments seem easier to
write and sometimes interviewees mentioned them as the way reviewers use to
avoid spending time to conduct good code reviews. An observation by a senior
developer, in the company for more than nine years, summarizes the opinions we
received from many interviewees: \quotation{Iíve seen quite a few code reviews
where someone commented on formatting while missing the fact that there were
security issues or data model issues.}

\subsection{Alternative Solutions}

``Alternative solutions'' regard changes and comments on improving the
submitted code by adopting an idea that leads to a better implementation. This
is one of the few motivations in which developers and managers do not agree.
While 147 (17\%) developers put this as the first motivation, 202 (23\%) as the
second, and 152 (17\%) as the third, only 4 managers (2\%) even mentioned it
(e.g., \quotation{Generate better ideas, alternative approaches} and
\quotation{Collective wisdom: Someone else on the project may have a better
idea to solve a problem}). The outcome of the interviews was similar to the
position of managers: Interviewees vaguely mentioned this motivation, and
mostly in terms of generic \quotation{better ways to do things.}

\subsection{Knowledge Transfer}

All the interviewees but one motivated their code reviews also from a learning,
or ``knowledge transfer,'' perspective. With the words of a senior developer:
\quotation{one of the things that should be happening with code reviews over
time is a distribution of knowledge. If you do a code review and did not learn
anything about the area and you still do not know anything about the area, then
that was not as good code review as it could have been.} Although we did not
include questions related to ``knowledge transfer'' in our interview guideline,
this topic kept emerging spontaneously from each meeting, thus underscoring its
value for practitioners.

Sometimes programmers told us that they follow code reviews explicitly for
learning purposes. For example, a tester explained: \quotation{[I read code
reviews because] from a code review you can learn about the different parts you
have to touch to implement a certain feature.}

According to interviewees, code review is a learning opportunity for both the
author of the change and the reviewers: There is a bidirectional knowledge
transfer about APIs usage, system design, best practices, team conventions,
\quotation{additional code tricks,} etc. Moreover code reviews are recognized
for educating new developers about code writing.

Managers included ``knowledge transfer'' as one of the reasons for code review,
although never as the top motivation. They mostly wrote about code review as an
education means by mentioning among the motivations for code review:
\quotation{developer education,} \quotation{education for junior developers who
are learning the codebase,} and \quotation{learning tool to teach more junior
team members.}

Programmers answering the survey declared ``knowledge transfer'' to be their
first motivation for code review in 73 cases (8\%), their second in 119 (14\%),
and their third in 141 (16\%).

\subsection{Team Awareness and Transparency}

During one of our observations, one developer was preparing a code review
submission as an author: He wanted other developers to \quotation{double check}
his changes before committing them to the repository. After preparing the code,
he specified the developers he wanted to review his code; he required not only
two specific people, but he also put a generic email distribution group as an
\quotation{optional} reviewer. When we inquired about this choice, he explained
us: \quotation{I am adding [this alias], so that everybody [in the team] is
notified about the change I want to do before I check it in.} In the subsequent
interviews, this concept of using an email list as optional reviewer, or
including specific optional reviewers exclusively for awareness emerged again
frequently, e.g., \quotation{Code reviews are good FYIs [for your
information].}

Managers often mentioned the concept of team awareness as a motivation for code
review, frequently justifying it with the notion of ``transparency:'' Not only
must the team be kept aware of the directions taken by the code, but also
nobody should be allowed to ``secretly'' make changes that might break the code
or alter functionalities.

The 873 programmers answering the survey ranked ``team awareness and
transparency'' very close to ``knowledge transfer.'' In fact, the two concepts
appeared logically related also in the interviews; for example one tester,
while reviewing some code said: \quotation{oh, this guy just implemented this
feature, and now let me back and use it somewhere else.} Showing that he both
learned about the new feature and he was now aware of the possibility to use it
in his own code. 75 (9\%) developers considered team awareness their first
motivation for code review, 108 (12\%) their second, and 149 (17\%) their
third.

Although team awareness and transparency emerged from our data as clearly
promoted by the code review process, academic research seems to have given
little attention to it.

\subsection{Share Code Ownership}

The concept of ``shared code ownership'' is closely related to ``team awareness
and transparency,'' but it has a stronger connotation toward active
collaboration and overlapping coding activities. Programmers and managers
believe that code review is not only an occasion to notify other team members
about incoming changes, but also a means to have more than one knowledgeable
person about specific parts of the codebase. A manager put the following as her
second motivation for code review: \quotation{Broaden knowledge \& understanding
of how specific features/areas are designed and implemented (e.g., grooming
``backup developers'' for areas where knowledge is too concentrated on one or
two expert developers.}

Moreover, both developers and managers have the opinion that practicing code
review also improves the personal perception of team members about shared code
ownership. On this note, a senior developer, with more than 30 years in the
software industry, explained: \quotation{In the past people did not use to do
code reviews and were very reluctant to put themselves in positions where they
were having other people critiquing their code. The fact that code reviews are
considered as a normal thing helps immensely with making people less protective
about their code.} Similarly a manager wrote us explaining that she deems code
reviews important because they \quotation{Dilute any ``rigid sense of
ownership'' that might develop over chunks of code.}

In the programmersí survey, 51 respondents (6\%) marked ``share code
ownership'' as their first motivation, 100 (11\%) as their second, and (10\%)
as their third.

\subsection{Summary}

In this section, we analyzed the motivations that developers and managers have
for doing code review. We abstracted them into a list, which we finally
included in the programmersí survey. Figure 3 reports the answers given to this
question: The black bar is the number of developers that put that row as their
top motivation, the gray bar is the number that put it as the second
motivation, etc. We have ordered the factors by giving 3 points for a first
motivation response, 2 points for a second motivation, etc. and then sorting by
the sum.

We discussed the five most prominent motivations, which show that ``finding
defects'' is the top motivation, although participants believe that code review
brings other benefits. The first two motivations were already popular in
research and their effectiveness have been evaluated in the context of code
inspections; on the contrary, the other motivations are still unexplored,
especially those regarding more ``social'' benefits on the team, such as shared
code ownership.

Although motivations are well defined, we still have to verify whether they
actually translate into real outcomes of a modern code review process.

### 2013-01-05 sections/implications.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Recommendations and Implications} \label{sec:implications}
%%%%%%%%%%%%%%%%%%%%%%

### 2013-01-05 sections/introduction.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%

Peer code review, a manual inspection of source code by developers other than
the author, is recognized as a valuable tool for reducing software defects and
improving the quality of software projects (Ackerman, Fowler and Ebenau 1984)
(Ackerman, Buchwald and Lewski 1989). In 1976, Fagan formalized a highly
structured process for code reviewing (Fagan 1976), based on line-by-line group
reviews, done in extended meetings -- code inspections. Over the years,
researchers provided evidence on code inspectionís benefits, especially in
terms of defect finding, but the cumbersome, time-consuming, and synchronous
nature of this approach hinders its universal adoption in practice (Shull
2008).

Nowadays, many organizations are adopting more lightweight code review
practices to limit the inefficiencies of inspections. In particular, there is a
clear trend toward the usage of tools specifically developed to support code
review (Rigby, Cleary, et al. 2012). In the context of this paper, we define
Modern Code Review, as review that is (1) informal (in contrast to
Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
nowadays, for example at companies such as Microsoft, Google (Kennedy 2006),
Facebook (Tsotsis 2011), and in other companies and OSS projects (Gerrit
(software) 2012).

This trend raises questions, such as: Can we apply the lessons learned from
previous research on code inspections to modern code reviews? What are the
expectations for code review nowadays? What are the actual outcomes of code
review? What challenges do people face in code review?

Answers to these questions can provide insight for both practitioners and
researchers.  Developers and other software project stakeholders can use
empirical evidence about expectations and outcomes to make informed decisions
about when to use code review and how it should fit into their development
process. Researchers can focus their attention on practitionersí challenges to
make code review more effective.

We present an in-depth study of practices in teams that use modern code review,
revealing what practitioners think, do, and achieve when it comes to modern
code review.

Since Microsoft is made up of many different teams working on very diverse
products, it gives the opportunity to study teams performing code review in
situ and understand their expectations, the benefits they derive from code
review, the needs they have, and the problems they face.

We set up our study as an explorative investigation. We started without a
priori hypotheses regarding how and why code review should be performed, with
the aim of discovering what developers and managers expect from code review,
how reviews are conducted in practice, and what the actual outcomes and
challenges are. To that end, we (1) observed 17 industrial developers
performing code review with various degrees of experience and seniority across
16 separate product teams with distinct reviewing cultures and policies; (2)
interviewed these developers using a semi-structured interviews; (3) manually
inspected and classified the content of 570 comments in discussions contained
within code reviews; and (4) surveyed 165 managers and 873 programmers.

Our results show that, although the top motivation driving code reviews is
still finding defects, the practice and the actual outcomes are less about
finding errors than expected: Defect related comments comprise a small
proportion and mainly cover small logical low-level issues. At the same time,
code review additionally provides a wide spectrum of benefits to software
teams, such as knowledge transfer, team awareness, and improved solutions to
problems. Moreover, we found that context and change understanding is the key
of any review. According to the outcomes they want to achieve, developers
employ many mechanisms to fulfill their understanding needs, most of which are
not currently met by any code review tool.

This paper makes the following contributions:
\begin{itemize}
  \item Characterizing the motivations of developers and managers for code review and compare with actual outcomes.
  \item Relating the outcomes to understanding needs and discuss how developers achieve such needs.
\end{itemize}

Based on our findings, we provide recommendations for practitioners and
implications for researchers as well as outline future avenues for research.

### 2013-01-05 sections/limitations.tex
moved sections into a separate dir

--- 2013-01-05 sections/introduction.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction} \label{sec:introduction}
+\section{Limitations} \label{sec:limitations}
 %%%%%%%%%%%%%%%%%%%%%%
-
-Peer code review, a manual inspection of source code by developers other than
-the author, is recognized as a valuable tool for reducing software defects and
-improving the quality of software projects (Ackerman, Fowler and Ebenau 1984)
-(Ackerman, Buchwald and Lewski 1989). In 1976, Fagan formalized a highly
-structured process for code reviewing (Fagan 1976), based on line-by-line group
-reviews, done in extended meetings -- code inspections. Over the years,
-researchers provided evidence on code inspectionís benefits, especially in
-terms of defect finding, but the cumbersome, time-consuming, and synchronous
-nature of this approach hinders its universal adoption in practice (Shull
-2008).
-
-Nowadays, many organizations are adopting more lightweight code review
-practices to limit the inefficiencies of inspections. In particular, there is a
-clear trend toward the usage of tools specifically developed to support code
-review (Rigby, Cleary, et al. 2012). In the context of this paper, we define
-Modern Code Review, as review that is (1) informal (in contrast to
-Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
-nowadays, for example at companies such as Microsoft, Google (Kennedy 2006),
-Facebook (Tsotsis 2011), and in other companies and OSS projects (Gerrit
-(software) 2012).
-
-This trend raises questions, such as: Can we apply the lessons learned from
-previous research on code inspections to modern code reviews? What are the
-expectations for code review nowadays? What are the actual outcomes of code
-review? What challenges do people face in code review?
-
-Answers to these questions can provide insight for both practitioners and
-researchers.  Developers and other software project stakeholders can use
-empirical evidence about expectations and outcomes to make informed decisions
-about when to use code review and how it should fit into their development
-process. Researchers can focus their attention on practitionersí challenges to
-make code review more effective.
-
-We present an in-depth study of practices in teams that use modern code review,
-revealing what practitioners think, do, and achieve when it comes to modern
-code review.
-
-Since Microsoft is made up of many different teams working on very diverse
-products, it gives the opportunity to study teams performing code review in
-situ and understand their expectations, the benefits they derive from code
-review, the needs they have, and the problems they face.
-
-We set up our study as an explorative investigation. We started without a
-priori hypotheses regarding how and why code review should be performed, with
-the aim of discovering what developers and managers expect from code review,
-how reviews are conducted in practice, and what the actual outcomes and
-challenges are. To that end, we (1) observed 17 industrial developers
-performing code review with various degrees of experience and seniority across
-16 separate product teams with distinct reviewing cultures and policies; (2)
-interviewed these developers using a semi-structured interviews; (3) manually
-inspected and classified the content of 570 comments in discussions contained
-within code reviews; and (4) surveyed 165 managers and 873 programmers.
-
-Our results show that, although the top motivation driving code reviews is
-still finding defects, the practice and the actual outcomes are less about
-finding errors than expected: Defect related comments comprise a small
-proportion and mainly cover small logical low-level issues. At the same time,
-code review additionally provides a wide spectrum of benefits to software
-teams, such as knowledge transfer, team awareness, and improved solutions to
-problems. Moreover, we found that context and change understanding is the key
-of any review. According to the outcomes they want to achieve, developers
-employ many mechanisms to fulfill their understanding needs, most of which are
-not currently met by any code review tool.
-
-This paper makes the following contributions:
-\begin{itemize}
-  \item Characterizing the motivations of developers and managers for code review and compare with actual outcomes.
-  \item Relating the outcomes to understanding needs and discuss how developers achieve such needs.
-\end{itemize}
-
-Based on our findings, we provide recommendations for practitioners and
-implications for researchers as well as outline future avenues for research.

### 2013-01-05 sections/methodology.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology} \label{sec:methodology}
%%%%%%%%%%%%%%%%%%%%%%

In this section we define the research questions, describe the research
settings, and outline our research method.

\subsection{Research Questions}

Our investigation of code review revolves around the following research
questions, which we iteratively refined during our initial in-field
observations and interviews:

\begin{enumerate}
  \item What are the motivations and expectations for modern code review? Do they change from managers to developers and testers?
  \item What are the actual outcomes of modern code review? Do they match the expectations?
  \item What are the main challenges experienced when performing modern code reviews relative to the expectations and outcomes?
\end{enumerate}

\subsection{Research Setting}

Our study took place with professional developers, testers, and managers.
Microsoft develops software in diverse domains, from high end server enterprise
data management solutions such as SQL Server to mobile phone applications and
smart phone apps to search engines. Each team has its own development culture
and code review policies. Over the past two years, a common tool for code
review at Microsoft has achieved wide-spread adoption. As it represents a
common and growing solution for code review (over 40,000 developers used it so
far), we focused on developers using this tool for code review -- CodeFlow.

CodeFlow is a collaborative code review tool that allows users to directly
annotate source code in its viewer and interact with review participants in a
live chat model. The functionality of CodeFlow is similar to other review tools
such Googleís Mondrian (Kennedy 2006), Facebookís Phabricator (Tsotsis 2011) or
open-source  Gerrit (Gerrit (software) 2012). Developers who want their code to
be reviewed create a package with the changed (new, deleted, and modified)
files, select the reviewers, write a message to describe the code review, and
submit everything to the CodeFlow service. CodeFlow then notifies the reviewers
about the incoming task via email.

Once reviewers open a CodeFlow review, they interact with it via a single
desktop window (Figure 1). On the top left (1), they see the list of files
changed in the current submission, plus a ``description.txt'' file, which
contains a textual explanation of the change, written by the author. On bottom
left, CodeFlow shows the list of reviewers and their status (2). We see that
Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
Alberto has reviewed and is waiting for the author to act, as the clock icon
suggests, while Nachi already signed off on the changes. CodeFlowís main view
(3) shows the diff-highlighted content of the file currently under review. Both
the reviewers and the author can highlight portions of the code and add
comments inline (4). These comments can start threads of discussion and are the
interaction points for the people involved in the review. Each user viewing the
same review in CodeFlow sees events as they happen.  Thus, if an author and
reviewer are working on the review at the same time, the communication is
synchronous and comment threads act similar to instant messaging. The comments
are persisted so that if they work at different times, the communication
becomes asynchronous. The bottom right pane (5) shows the summary of all the
comments in the review.

CodeFlow centralizes and records all the information on code reviews on a
central server. This provides an additional data source that we used to analyze
real code review comments without incurring the Hawthorne effect (Adair 1984).

\subsection{Research Method}

Our research method followed a mixed approach (Creswell 2009), depicted in
Figure 2, collecting data from different sources for triangulation: (1)
analysis of previous study, (2) observations and interviews with developers,
(3) card sort on interview data,  (4) card sort on code review comments, (5)
the creation of an affinity diagram, and (6) survey to managers and
programmers.

1. Analysis of previous study: Our research started with the analysis of a
study commissioned by Microsoft, between April and May 2012 carried out by an
external vendor. The study investigated how different product teams were using
CodeFlow. It consisted of structured interviews (lasting 30-50 minutes) to 23
people with different roles.

Most of the interview questions revolved around topics that are very specific
to tool usage, and were only tangentially related to this work. We found one
relevant as a starting point for our study: ``What do you hope to accomplish
when you submit a code review?'' We analyzed the transcript of this answer, for
each interview, through the process of coding (Berg 2004) (also used in
grounded theory (Adolph 2011)): breaking up the answers into smaller coherent
units (sentences or paragraphs) and adding codes to them. We organized codes
into concepts, which in turn were grouped into more abstract categories.

From this analysis, four motivations emerged for code review: finding defects,
maintaining team awareness, improving code quality, and assessing the
high-level design. We used them to draw an initial guideline for our
interviews.

2. Observations and interviews with developers: Next, we conducted a series of
one-to-one meetings with developers who use CodeFlow, each taking 40-60
minutes.

We contacted 100 random candidates who signed-off between 50 and 250 code
reviews since the CodeFlow release and sampled across different product teams
to address our research questions from a multi-point perspective. We wrote
developers who used CodeFlow in the past and asked them to contact us, giving
us 30 minute notice when they received their next review task so that we could
observe.  The respondents that we interviewed comprised five developers, four
senior developers, six testers, one senior tester, and one software architect.
Their time in the company ranged from 18 months to almost 10 years, with a
median of five years.

Each meeting was comprised of two parts: In the first part, we observed them
performing the code review that they had been assigned. To minimize
invasiveness we used only one observer and to encourage the participant to
narrate their work, we asked the participants to think of us as a newcomer to
the team. In this way, most developers thought aloud without need of prompting.
With consent, we recorded the audio, assuring the participants of anonymity.
Since we, as observers, have backgrounds in software development and practices
at Microsoft, we were able to understand most of the work and where and how
information was obtained without inquiry.

The second part of the meeting was a semi-structured interview (Lindlof and
Taylor 2002). Semi-structured interviews make use of an interview guide that
contains general groupings of topics and questions rather than a pre-determined
exact set and order of questions.  They are often used in an exploratory
context to ``find out what is happening [and] to seek new insights'' (Weiss
1995). The guideline was iteratively refined after each interview, in
particular when developers started providing answers very similar to the
earlier ones, thus reaching a saturation effect.

Observations also reached a saturation point, thus providing insights very similar to the earlier ones. For this, after the first 5-6 observations, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.

The audio of each interview was then transcribed and broken up into smaller
coherent units for subsequent analysis.

3. Card sort (meetings): To group codes that emerged from interviews and
observations into categories, we conducted a card sort. Card sorting is a
sorting technique that is widely used in information architecture to create
mental models and derive taxonomies from input data (Barker 2005). In our case
it helped to organize the codes into hierarchies to deduce a higher level of
abstraction and identify common themes. A card sort involves three phases: In
the (1) preparation phase, participants of the card sort are selected and the
cards are created; in the (2) execution phase, cards are sorted into meaningful
groups with a descriptive title; and in the (3) analysis phase, abstract
hierarchies are formed to deduce general categories.

We applied an open card sort: There were no predefined groups. Instead, the
groups emerged and evolved during the sorting process. In contrast, a closed
card sort has predefined groups and is typically applied when themes are known
in advance, which was not the case for our study.

The first author of this paper created all of the cards, from the 1,047
coherent units in the interviews. Throughout our further analysis other
researchers (the second author and external people) were involved in developing
categories and assigning cards to categories, so as to strengthen the validity
of the result. The first author played a special role of ensuring that the
context of each question was appropriately considered in the categorization,
and creating the initial categories. To ensure the integrity of our categories,
the cards were sorted by the first author several times to identify initial
themes. Next, all researchers reviewed and agreed on the final set of
categories.

4. Card sort (code review comments): The same method was applied to group code
review comments into categories: We randomly sampled 200 threads with at least
two comments (e.g., Point 4 of Figure 2), from the entire dataset of CodeFlow
reviews, which embeds data from dozens of independent software products at
Microsoft. We printed one card for each comment (along with the entire
discussion thread to give the context), totaling 570 cards, and conducted a
card sort, as performed for the interviews, to identify common themes.

5. Affinity Diagram: We used an affinity diagram to organize the categories
that emerged from the card sort. This tool allows large numbers of ideas to be
sorted into groups for review and analysis (Stuart J. Janis 2000). We used it
to generate an overview of the topics that emerged from the card sort, in order
to connect the related concepts and derive the main themes. For generating the
affinity diagram, we followed the five canonical steps: we (1) recorded the
categories on post-it-notes, (2) spread them onto a wall, (3) sorted the
categories based on discussions, until all are sorted and all participants
agreed, (4) named each group with a description, and (5) captured and discussed
the themes. 6. Surveys: The final step of our study was aimed at validating the
concepts that emerged from the previous phases. Towards this goal, we created
two surveys to reach a significant number of participants and to challenge our
conclusions (The full surveys are available as a technical report (Bacchelli
and Bird 2012)). For the design of the surveys, we followed Kitchenham and
Pfleegerís guidelines for personal opinion surveys (Kitchenham and Pfleeger
2007). Both surveys were anonymous to increase response rates (Tyagi 1989).

We sent the first survey to a cross section of managers.  We considered
managers for which at least half of their team performed code reviews regularly
(on average, one per week or more) and sampled along two dimensions.  The first
dimension was whether or not the manager had participated in a code review
himself since the beginning of the year and the second dimension was whether
the manager managed a single team or multiple teams (a manager of managers).
Thus, we had one sample of first level managers who participated in review,
another sample of second level managers who participated in reviews, etc.  The
first survey was a short survey comprising 6 questions (all optional), which we
sent to 600 managers that had at least ten direct or indirect reporting
developers who used CodeFlow in the past. The central focus was the open
question asking to enumerate the main motivations for doing code reviews in
their team. We received 165 answers (28\% response rate), which we analyzed
before devising the second survey

The second survey comprised 18 questions, mostly closed with multiple choice
answers, and was sent to 2,000 randomly chosen developers who signed off on
average at least one code review per week since the beginning of the year. We
used the time frame of January to June of 2012 to minimize the amount of
organizational churn during the time period and identify employeesí activity in
their current role and team.  We received 873 answers (44\% response rate). Both
response rates were  high, as other online surveys in software engineering have
reported response rates ranging from 14\% to 20\% (Punter, et al. 2003).

### 2013-01-05 sections/related_work.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work} \label{sec:related_work}
%%%%%%%%%%%%%%%%%%%%%%

Previous studies exist that have examined the practices of code inspection and
code review.  Stein et al. conducted a study focusing specifically on
distributed, asynchronous code inspections [17]. The study included evaluation
of a tool that allowed for identification and sharing of code faults or
defects. Participants at separated locations can then discuss faults via the
tool. Laitenburger conducted a survey of code inspection methods, and presented
a taxonomy of code inspection techniques (Laitenberger 2002). Johnson conducted
an investigation into code review practices in open source development and the
effect they have on choices made by software project managers (Johnson 2006).

Porter et al. (Porter, Siy and Votta 1996) reported on a review of studies on
code inspection in 1995 that examined the effects of factors such as team size,
type of review, and number of sessions on code inspections.  They also assessed
costs and benefits across a number of studies.  These studies differ from ours
in that they were not tool-based and were the majority involved planned
meetings to discuss the code.

However, prior research also sheds light on why review today is more often
tool-based, informal, and often asynchronous. The current state of code review
might be due to the time required for more formal inspections.  Votta found
that 20\% of the interval in a ``traditional inspection'' is wasted due to
scheduling (Votta 1993). The ICICLE tool (Brothers, Sembugamoorthy and Muller
1990), or ``Intelligent Code Inspection in a C Language Environment,'' was
developed after researchers at Bellcore observed how much time and work was
expended before and during formal code inspections. Many of todayís review
tools are based on ideas that originated in ICICLE.  Other similar tools have
been developed in an effort to reduce time for inspection and allow
asynchronous work on reviews.   Examples include CAIS  (Mashayekhi, Feulner and
Riedl 1994) and Scrutiny (Gintell, et al. 1993).

More recently, Rigby has done extensive work examining code review practices in
open source software development (Rigby, Cleary, et al. 2012).  For example in
a study of practices in the Apache project (Rigby, German and Storey, Open
source software peer review practices: a case study of the apache server 2008)
they data-mined the email archives and found that reviews were typically small
and frequent, and that the contributions to a review were often brief and
independent from one another.

Sutherland and Venolia conducted a study at Microsoft regarding using code
review data for later information needs (Sutherland and Venolia 2009). They
hypothesized that the knowledge exchanged during code reviews could be of great
value to engineers later trying to understand or modify the discussed code.
They found that ``the meat of the code review dialog, no matter what medium, is
the articulation of design rationale'' and, thus, ``code reviews are an enticing
opportunity for capturing design rationale.''

When studying developer work habits, Latoza et al. found that many problems
encountered by developers were related to understanding the rationale behind
code changes and gathering knowledge from other members of their team (LaToza,
Venolia and DeLine 2006).

### 2013-01-06 sections/introduction.tex
done up to section IV (partially included)

@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
@@
 
+%   \vspace{-0.5em}
+
 Peer code review, a manual inspection of source code by developers other than
 the author, is recognized as a valuable tool for reducing software defects and
-improving the quality of software projects (Ackerman, Fowler and Ebenau 1984)
-(Ackerman, Buchwald and Lewski 1989). In 1976, Fagan formalized a highly
-structured process for code reviewing (Fagan 1976), based on line-by-line group
-reviews, done in extended meetings -- code inspections. Over the years,
-researchers provided evidence on code inspectionís benefits, especially in
+improving the quality of software projects~\cite{ackerman1984software, ackerman1989software}.
+In 1976, Fagan formalized a highly
+structured process for code reviewing~\cite{fagan1976design}, based on line-by-line group
+reviews, done in extended meetings---\emph{code inspections}. Over the years,
+researchers provided evidence on code inspection's benefits, especially in
 terms of defect finding, but the cumbersome, time-consuming, and synchronous
-nature of this approach hinders its universal adoption in practice (Shull
-2008).
+nature of this approach hinders its universal adoption in practice~\cite{shull2008inspecting}. %(Shull 2008).
 
@@
 practices to limit the inefficiencies of inspections. In particular, there is a
-clear trend toward the usage of tools specifically developed to support code
-review (Rigby, Cleary, et al. 2012). In the context of this paper, we define
-Modern Code Review, as review that is (1) informal (in contrast to
+clear trend toward the usage of tools developed to support code
+review~\cite{rigby2012open}. In the context of this paper, we define
+\emph{Modern Code Review}, as review that is (1) informal (in contrast to
 Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
-nowadays, for example at companies such as Microsoft, Google (Kennedy 2006),
-Facebook (Tsotsis 2011), and in other companies and OSS projects (Gerrit
-(software) 2012).
+nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
+Facebook~\cite{tsotsis2011online}, and in other companies and OSS projects~\cite{gerrit2012online}.
 
@@
 about when to use code review and how it should fit into their development
-process. Researchers can focus their attention on practitionersí challenges to
+process. Researchers can focus their attention on practitioners' challenges to
 make code review more effective.
@@
 Since Microsoft is made up of many different teams working on very diverse
-products, it gives the opportunity to study teams performing code review in
-situ and understand their expectations, the benefits they derive from code
+products, it gives the opportunity to study teams performing code review \emph{in
+situ} and understand their expectations, the benefits they derive from code
 review, the needs they have, and the problems they face.
 
-We set up our study as an explorative investigation. We started without a
-priori hypotheses regarding how and why code review should be performed, with
+We set up our study as an explorative investigation. We started without \emph{a
+priori} hypotheses regarding how and why code review should be performed, with
 the aim of discovering what developers and managers expect from code review,

### 2013-01-06 sections/methodology.tex
done up to section IV (partially included)

@@
 common and growing solution for code review (over 40,000 developers used it so
-far), we focused on developers using this tool for code review -- CodeFlow.
+far), we focused on developers using this tool for code review---\emph{CodeFlow}.
 
@@
 live chat model. The functionality of CodeFlow is similar to other review tools
-such Googleís Mondrian (Kennedy 2006), Facebookís Phabricator (Tsotsis 2011) or
-open-source  Gerrit (Gerrit (software) 2012). Developers who want their code to
+such Google's Mondrian~\cite{kennedy2006online}, Facebook's Phabricator~\cite{tsotsis2011online} or
+open-source Gerrit~\cite{gerrit2012online}. Developers who want their code to
 be reviewed create a package with the changed (new, deleted, and modified)
@@
 Once reviewers open a CodeFlow review, they interact with it via a single
-desktop window (Figure 1). On the top left (1), they see the list of files
+desktop window (Figure~\ref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
 changed in the current submission, plus a ``description.txt'' file, which
 contains a textual explanation of the change, written by the author. On bottom
-left, CodeFlow shows the list of reviewers and their status (2). We see that
+left, CodeFlow shows the list of reviewers and their status (\textbf{2}). We see that
 Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
 Alberto has reviewed and is waiting for the author to act, as the clock icon
-suggests, while Nachi already signed off on the changes. CodeFlowís main view
-(3) shows the diff-highlighted content of the file currently under review. Both
+suggests, while Nachi already signed off on the changes. CodeFlow's main view
+(\textbf{3}) shows the diff-highlighted content of the file currently under review. Both
 the reviewers and the author can highlight portions of the code and add
-comments inline (4). These comments can start threads of discussion and are the
+comments inline (\textbf{4}). These comments can start threads of discussion and are the
 interaction points for the people involved in the review. Each user viewing the
@@
 are persisted so that if they work at different times, the communication
-becomes asynchronous. The bottom right pane (5) shows the summary of all the
+becomes asynchronous. The bottom right pane (\textbf{5}) shows the summary of all the
 comments in the review.
@@
 central server. This provides an additional data source that we used to analyze
-real code review comments without incurring the Hawthorne effect (Adair 1984).
+real code review comments without incurring the Hawthorne effect~\cite{adair1984hawthorne}.
+
+\begin{figure}[t] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=0.98\columnwidth]{figs/codeflow.pdf}
+   \caption{Screenshot of CodeFlow, the dominant code review tool used by developers at Microsoft.}
+   \label{fig:codeflow:screenshot}
+      \vspace{-1.5em}
+\end{figure}
 
@@
 
-Our research method followed a mixed approach (Creswell 2009), depicted in
-Figure 2, collecting data from different sources for triangulation: (1)
-analysis of previous study, (2) observations and interviews with developers,
-(3) card sort on interview data,  (4) card sort on code review comments, (5)
-the creation of an affinity diagram, and (6) survey to managers and
+\begin{figure*}[t] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=1.3\columnwidth]{figs/methodology.pdf}
+   \caption{The mixed approach research method applied.}
+   \label{fig:research-method}
+   \vspace{-1.5em}
+\end{figure*}
+
+Our research method followed a mixed approach~\cite{creswell2009research}, depicted in
+Figure~\ref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
+analysis of previous study, (\textbf{2}) observations and interviews with developers,
+(\textbf{3}) card sort on interview data,  (\textbf{4}) card sort on code review comments, (\textbf{5})
+the creation of an affinity diagram, and (\textbf{6}) survey to managers and
 programmers.
 
-1. Analysis of previous study: Our research started with the analysis of a
+\textbf{1. Analysis of previous study:} Our research started with the analysis of a
 study commissioned by Microsoft, between April and May 2012 carried out by an
@@
 to tool usage, and were only tangentially related to this work. We found one
-relevant as a starting point for our study: ``What do you hope to accomplish
-when you submit a code review?'' We analyzed the transcript of this answer, for
-each interview, through the process of coding (Berg 2004) (also used in
-grounded theory (Adolph 2011)): breaking up the answers into smaller coherent
-units (sentences or paragraphs) and adding codes to them. We organized codes
-into concepts, which in turn were grouped into more abstract categories.
+relevant as a starting point for our study: ``\emph{What do you hope to accomplish
+when you submit a code review?}'' We analyzed the transcript of this answer, for
+each interview, through the process of \emph{coding}~\cite{berg2004qualitative} (also used in
+\emph{grounded theory}~\cite{adolph2011using}): breaking up the answers into smaller coherent
+units (sentences or paragraphs) and adding \emph{codes} to them. We organized codes
+into \emph{concepts}, which in turn were grouped into more abstract \emph{categories}.
 
@@
 
-2. Observations and interviews with developers: Next, we conducted a series of
+\textbf{2. Observations and interviews with developers:} Next, we conducted a series of
 one-to-one meetings with developers who use CodeFlow, each taking 40-60
@@
 reviews since the CodeFlow release and sampled across different product teams
-to address our research questions from a multi-point perspective. We wrote
+to address our research questions from a \emph{multi-point} perspective. We wrote
 developers who used CodeFlow in the past and asked them to contact us, giving
@@
 
-The second part of the meeting was a semi-structured interview (Lindlof and
-Taylor 2002). Semi-structured interviews make use of an interview guide that
+The second part of the meeting was a \emph{semi-structured} interview~\cite{taylor2010qualitative}. Semi-structured interviews make use of an \emph{interview guide} that
 contains general groupings of topics and questions rather than a pre-determined
 exact set and order of questions.  They are often used in an exploratory
-context to ``find out what is happening [and] to seek new insights'' (Weiss
-1995). The guideline was iteratively refined after each interview, in
+context to ``find out what is happening [and] to seek new insights''~\cite{weiss1995learning}. 
+The guideline was iteratively refined after each interview, in
 particular when developers started providing answers very similar to the
@@
 
-Observations also reached a saturation point, thus providing insights very similar to the earlier ones.
+Observations also reached a \emph{saturation} point, thus providing insights very similar to the earlier ones.
 For this, after the first 5-6 observations, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.
@@
 
-3. Card sort (meetings): To group codes that emerged from interviews and
-observations into categories, we conducted a card sort. Card sorting is a
+\textbf{3. Card sort (meetings):} To group codes that emerged from interviews and
+observations into categories, we conducted a \emph{card sort}. Card sorting is a
 sorting technique that is widely used in information architecture to create
-mental models and derive taxonomies from input data (Barker 2005). In our case
+mental models and derive taxonomies from input data~\cite{barker2005online}. In our case
 it helped to organize the codes into hierarchies to deduce a higher level of
 abstraction and identify common themes. A card sort involves three phases: In
-the (1) preparation phase, participants of the card sort are selected and the
-cards are created; in the (2) execution phase, cards are sorted into meaningful
-groups with a descriptive title; and in the (3) analysis phase, abstract
+the (1) \emph{preparation phase}, participants of the card sort are selected and the
+cards are created; in the (2) \emph{execution phase}, cards are sorted into meaningful
+groups with a descriptive title; and in the (3) \emph{analysis phase}, abstract
 hierarchies are formed to deduce general categories.
 
-We applied an open card sort: There were no predefined groups. Instead, the
+We applied an \emph{open card sort}: There were no predefined groups. Instead, the
 groups emerged and evolved during the sorting process. In contrast, a closed
@@
 
-4. Card sort (code review comments): The same method was applied to group code
+\textbf{4. Card sort (code review comments):} The same method was applied to group code
 review comments into categories: We randomly sampled 200 threads with at least
-two comments (e.g., Point 4 of Figure 2), from the entire dataset of CodeFlow
+two comments (e.g., Point 4 of Figure~\ref{fig:research-method}), from the entire dataset of CodeFlow
 reviews, which embeds data from dozens of independent software products at
@@
 
-5. Affinity Diagram: We used an affinity diagram to organize the categories
+\textbf{5. Affinity Diagram:} We used an \emph{affinity diagram} to organize the categories
 that emerged from the card sort. This tool allows large numbers of ideas to be
-sorted into groups for review and analysis (Stuart J. Janis 2000). We used it
+sorted into groups for review and analysis~\cite{shade2000improving}. We used it
 to generate an overview of the topics that emerged from the card sort, in order
@@
 agreed, (4) named each group with a description, and (5) captured and discussed
-the themes. 6. Surveys: The final step of our study was aimed at validating the
+the themes.
+
+\textbf{6. Surveys:} The final step of our study was aimed at validating the
 concepts that emerged from the previous phases. Towards this goal, we created
 two surveys to reach a significant number of participants and to challenge our
-conclusions (The full surveys are available as a technical report (Bacchelli
-and Bird 2012)). For the design of the surveys, we followed Kitchenham and
-Pfleegerís guidelines for personal opinion surveys (Kitchenham and Pfleeger
-2007). Both surveys were anonymous to increase response rates (Tyagi 1989).
+conclusions (The full surveys are available as a technical report~\cite{bacchelli2012appendix}). 
+For the design of the surveys, we followed Kitchenham and
+Pfleeger's guidelines for personal opinion surveys~\cite{kitchenham2008personal}. 
+Both surveys were anonymous to increase response rates~\cite{tyagi1989effects}.
 
@@
 Thus, we had one sample of first level managers who participated in review,
-another sample of second level managers who participated in reviews, etc.  The
+another sample of second level managers who participated in reviews, \etc.  The
 first survey was a short survey comprising 6 questions (all optional), which we
@@
 their team. We received 165 answers (28\% response rate), which we analyzed
-before devising the second survey
+before devising the second survey.
 
@@
 used the time frame of January to June of 2012 to minimize the amount of
-organizational churn during the time period and identify employeesí activity in
+organizational churn during the time period and identify employees' activity in
 their current role and team.  We received 873 answers (44\% response rate). Both
 response rates were  high, as other online surveys in software engineering have
-reported response rates ranging from 14\% to 20\% (Punter, et al. 2003).
+reported response rates ranging from 14\% to 20\%~\cite{punter2003conducting}.

### 2013-01-06 sections/related_work.tex
done up to section IV (partially included)

@@
 Previous studies exist that have examined the practices of code inspection and
-code review.  Stein et al. conducted a study focusing specifically on
-distributed, asynchronous code inspections [17]. The study included evaluation
+code review.  Stein \etal conducted a study focusing specifically on
+distributed, asynchronous code inspections~\cite{stein1997casestudy}. The study included evaluation
 of a tool that allowed for identification and sharing of code faults or
@@
 tool. Laitenburger conducted a survey of code inspection methods, and presented
-a taxonomy of code inspection techniques (Laitenberger 2002). Johnson conducted
-an investigation into code review practices in open source development and the
-effect they have on choices made by software project managers (Johnson 2006).
+a taxonomy of code inspection techniques~\cite{laitenberger2002survey}. Johnson conducted
+an investigation into code review practices in open source development and their
+effect on choices made by software project managers~\cite{johnson2006collaboration}.
 
-Porter et al. (Porter, Siy and Votta 1996) reported on a review of studies on
+Porter \etal~\cite{porter1996review} reported on a review of studies on
 code inspection in 1995 that examined the effects of factors such as team size,
@@
 that 20\% of the interval in a ``traditional inspection'' is wasted due to
-scheduling (Votta 1993). The ICICLE tool (Brothers, Sembugamoorthy and Muller
-1990), or ``Intelligent Code Inspection in a C Language Environment,'' was
+scheduling~\cite{votta1993does}. The ICICLE tool~\cite{brothers1990icicle}, or ``Intelligent Code Inspection in a C Language Environment,'' was
 developed after researchers at Bellcore observed how much time and work was
@@
 been developed in an effort to reduce time for inspection and allow
-asynchronous work on reviews.   Examples include CAIS  (Mashayekhi, Feulner and
-Riedl 1994) and Scrutiny (Gintell, et al. 1993).
+asynchronous work on reviews.   Examples include CAIS~\cite{mashayekhi1994cais} 
+and Scrutiny~\cite{gintell1993scrutiny}.
 
 More recently, Rigby has done extensive work examining code review practices in
-open source software development (Rigby, Cleary, et al. 2012).  For example in
-a study of practices in the Apache project (Rigby, German and Storey, Open
-source software peer review practices: a case study of the apache server 2008)
+open source software development~\cite{rigby2012open}.  For example in
+a study of practices in the Apache project~\cite{rigby2008open}
 they data-mined the email archives and found that reviews were typically small
@@
 Sutherland and Venolia conducted a study at Microsoft regarding using code
-review data for later information needs (Sutherland and Venolia 2009). They
+review data for later information needs~\cite{sutherland2009can}. They
 hypothesized that the knowledge exchanged during code reviews could be of great
@@
 
-When studying developer work habits, Latoza et al. found that many problems
+When studying developer work habits, Latoza \etal found that many problems
 encountered by developers were related to understanding the rationale behind
-code changes and gathering knowledge from other members of their team (LaToza,
-Venolia and DeLine 2006).
+code changes and gathering knowledge from other members of their team~\cite{latoza2006maintaining}.

### 2013-01-06 sections/challenges.tex
done up to section IV (partially included)
done with section 6

@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction} \label{sec:introduction}
+\section{What Are The Challenges of Code Review?} \label{sec:challenges}
 %%%%%%%%%%%%%%%%%%%%%%
+
+Our third research question seeks to understand the main challenges faced by the reviewers when performing modern code reviews, also with respect to the expected outcomes.
+We also seek to uncover the reason behind the mismatch between expectations and actual outcomes on finding defects in code reviews.
+
+\subsection{Code Review is Understanding}
+
+Even though we did not ask any specific question concerning understanding, the theme emerged clearly from our interviews.
+Many interviewees eventually acknowledged that understanding is their main challenge when doing code reviews.
+For example, a senior developer autonomously explained to us: \quotation{the most difficult thing when doing a code review is understanding the reason of the change;} a tester, in the same vein: \quotation{the biggest information need in code review: what instigated the change;} and another senior developer: \quotation{in a successful code review submission the author is sure that his peers understand and approve the change.} Although the textual description should help reviewers understanding, some developers do not find it useful: \quotation{people can say they are doing one thing, while they are doing many more of them,} or \quotation{the description is not enough;} in general, developers seem to confirm that \quotation{not knowing files (or {\normalfont [dealing with]} new ones) is a major reason for not understanding a change.}
+
+From interviews, no other code review challenge emerged as clearly as understanding the submitted change.
+Even though scheduling and time issues also appeared challenging, we could always trace them back to the first challenge—through the words of a tester: \quotation{understanding the code takes most of the reviewing time.} On the same note, in the code review comments we analyzed, the second most frequent category concerns understanding.
+This category includes clarification questions and doubts raised by the reviewers who want to grasp the rationale of the changes done on the code, and the corresponding clarification answers.
+This is also in line with the evidence delivered by Sutherland \& Venolia on the relevance of rationale articulation in reviews~\cite{sutherland2009can}.
+
+Do understanding needs change with the expected outcome of code review?
+We included a question in the programmers’ survey to know how much understanding they needed to achieve each of the motivations listed in \figref{XXX3}.
+The outcome of the question is summarized in \figref{XXX5}.
+The respondents could answer with a four values Likert's scale, by selecting the understanding of the change they felt was required to achieve the specific outcome.
+The most difficult task from the understanding perspective is ``finding defects,'' immediately followed by ``alternative solutions.'' Both clearly stand out from the other items.
+The gap in understanding needs between ``finding defects'' and ``code improvement'' seems to corroborate our hypothesis that the difference in the number of comments about these two items in review comments is mostly due to understanding issues.
+Thus, if managers and developers want code review to match their need for ``finding defects,'' context and change understanding must be improved.
+
+\subsection{Code Review is Understanding}
+
+By observing developers performing code reviews, we noticed that some started code reviews by thoroughly reading the accompanying textual description, while others went directly to a specific changed file.
+In the first group, the time required for putting the first review comments and understanding the change rationale was noticeably longer, and some of the comments were asking to clarify the reasons for a change.
+To better comprehend this situation, we included in our interview guideline a question about how the interviewees start code reviews.
+Participants explained that when they own or are very familiar with the files being changed, they have a better context and it is easier for them to understand the change submitted: \quotation{when doing code review I start with things I am familiar with, so it is easier to see what is going on.} When they are file owners, they often do not even need to read the description, but they \quotation{go directly to the files they own.} On the contrary, when they do not own files, or have to review new files, they need more information and try to get it from the description, which is deemed good when it states \quotation{what was changed and why.}
+
+To better understand this aspect we included two questions in the programmers' survey to know \begin{inparaenum}[(1)]
+\item whether it takes longer to review files they are not familiar with, and why; and 
+\item whether reviewers familiar with the changed files give different feedback, and how. \end{inparaenum}
+
+Most of the respondents (798, 91\%) answered positively to the first question, motivating it with the fact that it takes time to familiarize with the code and \quotation{learn enough about the files being modified to understand their purpose, invariants, APIs, etc.,} because \quotation{big-picture impact analysis requires contextual understanding.
+When reviewing a small, unfamiliar change, it is often necessary to read through much more code than that being reviewed.} The comment of a developer anticipates the answer to the second question: \quotation{It takes a lot longer to understand unknown code, but even then understanding isn't very deep.
+With code I am familiar with I have more to say.
+I know what to say faster.
+What I have to say is deeper.
+And I can be more insistent on it.} In fact, the answer to the second question is positive in 716~(82\%) cases.
+The main difference with file owner comments is that they are substantially deeper, more detailed and insightful.
+A respondent explained: \quotation{Comments reflect their deeper understanding -- more likely to find subtle defects, feedback is more conceptual (better ideas, approaches) instead of superficial (naming, mechanical style, etc.)} another tried to boldly summarize the concept: \quotation{Difference between algorithmic analysis and comments on coding style.
+The difference is big.}
+In fact, when the context is clear and understanding is very high, as in the case when the reviewer is the owner of changed files, code review authors receive comments that explore \quotation{deeper details,} are \quotation{more directed} and \quotation{more actionable and pertinent,} and find \quotation{more subtle issues.}
+
+\subsection{Dealing with Understanding Needs}
+
+From the interviews, we found that, in the current situation, reviewers try different paths to understand the context and the changes: They read the change description, try to run the changed code, send emails for understanding high level details about the review, and often (from 20\% to 40\% of the times) even go to talk in person to have a \quotation{higher communication bandwidth} for asking clarifications to the author.
+All code review tools that we see in practice today deliver only basic support for the understanding needs of reviewers -- providing features such as diffing capabilities, inline commenting, or syntax highlighting, which are limited when dealing with complex code understanding.

### 2013-01-06 sections/expectations.tex
done up to section IV (partially included)
fixed section title
done with section 6

@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Why Do Programmers Do Code Reviews} \label{sec:expectations}
+\section{Why Do Programmers Do Code Reviews?} \label{sec:expectations}
 %%%%%%%%%%%%%%%%%%%%%%
@@
 beneficial influences not only on the code, but also for the team and the
-entire development process. In this vein, one senior developerís comment
-summarized many of the responses: \quotation{[code review] also has several
-beneficial influences: (1) makes people less protective about their code, (2)
-gives another person insight into the code, so there is (3) better sharing of
-information across the team, (4) helps support coding conventions on the team,
-and [...] (5) helps improving the overall process and quality of code.}
+entire development process. In this vein, one senior developer's comment
+summarized many of the responses: \quotation{{\normalfont[code review]} also has several
+beneficial influences: \begin{inparaenum}[(1)]
+\item makes people less protective about their code, 
+\item gives another person insight into the code, so there is 
+\item better sharing of information across the team, 
+\item helps support coding conventions on the team,
+and {\normalfont[...]}
+\item helps improving the overall process and quality of code.
+\end{inparaenum}}
 
@@
 responses to create a comprehensive list of high-level motivations. We included
-this list in the developersí survey and asked them to rank the top three main
+this list in the developers' survey and asked them to rank the top three main
 reasons that described why they do code reviews.
@@
 doing code reviews; for 44\% of the managers, it is the top reason. Managers
-considered defects to be both low level issues (e.g., \quotation{correct logic
-is in place}) and high level concerns (e.g., \quotation{catch errors in
+considered defects to be both low level issues (\eg \quotation{correct logic
+is in place}) and high level concerns (\eg \quotation{catch errors in
 design}). Concerning surveyed developers/testers, ``finding defects'' is the
@@
 This is in-line with the reason why code inspections were devised in the first
-place: reducing software defects (Ackerman, Fowler and Ebenau 1984).
+place: reducing software defects~\cite{ackerman1984software}.
 
@@
 Code improvements are comments or changes about code in terms of readability,
-commenting, consistency, dead code removal, etc., but do not involve
+commenting, consistency, dead code removal, \etc, but do not involve
 correctness or defects.
@@
 (31\%). One manager wrote how code review in her view is a
-\quotation{discipline of explaining your code to your peers [that] drives a
+\quotation{discipline of explaining your code to your peers {\normalfont[that]} drives a
 higher standard of coding. I think the process is even more important than the
@@
 developer, in the company for more than nine years, summarizes the opinions we
-received from many interviewees: \quotation{Iíve seen quite a few code reviews
+received from many interviewees: \quotation{I've seen quite a few code reviews
 where someone commented on formatting while missing the fact that there were
@@
 is one of the few motivations in which developers and managers do not agree.
-While 147 (17\%) developers put this as the first motivation, 202 (23\%) as the
-second, and 152 (17\%) as the third, only 4 managers (2\%) even mentioned it
-(e.g., \quotation{Generate better ideas, alternative approaches} and
+While 147~(17\%) developers put this as the first motivation, 202~(23\%) as the
+second, and 152~(17\%) as the third, only 4~(2\%) managers even mentioned it
+(\eg \quotation{Generate better ideas, alternative approaches} and
 \quotation{Collective wisdom: Someone else on the project may have a better
@@
 Sometimes programmers told us that they follow code reviews explicitly for
-learning purposes. For example, a tester explained: \quotation{[I read code
-reviews because] from a code review you can learn about the different parts you
+learning purposes. For example, a tester explained: \quotation{{\normalfont[I read code
+reviews because]} from a code review you can learn about the different parts you
 have to touch to implement a certain feature.}
@@
 transfer about APIs usage, system design, best practices, team conventions,
-\quotation{additional code tricks,} etc. Moreover code reviews are recognized
+\quotation{additional code tricks,} \etc Moreover code reviews are recognized
 for educating new developers about code writing.
@@
 although never as the top motivation. They mostly wrote about code review as an
-education means by mentioning among the motivations for code review:
+education means by mentioning among the motivations:
 \quotation{developer education,} \quotation{education for junior developers who
@@
 Programmers answering the survey declared ``knowledge transfer'' to be their
-first motivation for code review in 73 cases (8\%), their second in 119 (14\%),
-and their third in 141 (16\%).
+first motivation for code review in 73~(8\%) cases, their second in 119~(14\%),
+and their third in 141~(16\%).
 
@@
 \quotation{optional} reviewer. When we inquired about this choice, he explained
-us: \quotation{I am adding [this alias], so that everybody [in the team] is
+us: \quotation{I am adding {\normalfont[this alias]}, so that everybody {\normalfont[in the team]} is
 notified about the change I want to do before I check it in.} In the subsequent
@@
 including specific optional reviewers exclusively for awareness emerged again
-frequently, e.g., \quotation{Code reviews are good FYIs [for your
-information].}
+frequently, \eg \quotation{Code reviews are good FYIs {\normalfont[for your
+information]}.}
 
@@
 learned about the new feature and he was now aware of the possibility to use it
-in his own code. 75 (9\%) developers considered team awareness their first
-motivation for code review, 108 (12\%) their second, and 149 (17\%) their
+in his own code. 75~(9\%) developers considered team awareness their first
+motivation for code review, 108~(12\%) their second, and 149~(17\%) their
 third.
@@
 ``backup developers'' for areas where knowledge is too concentrated on one or
-two expert developers.}
+two expert developers).}
 
@@
 
-In the programmersí survey, 51 respondents (6\%) marked ``share code
-ownership'' as their first motivation, 100 (11\%) as their second, and (10\%)
+In the programmers' survey, 51~(6\%) respondents marked ``share code
+ownership'' as their first motivation, 100~(11\%) as their second, and 91~(10\%)
 as their third.
@@
 for doing code review. We abstracted them into a list, which we finally
-included in the programmersí survey. Figure 3 reports the answers given to this
+included in the programmers' survey. Figure 3 reports the answers given to this
 question: The black bar is the number of developers that put that row as their
 top motivation, the gray bar is the number that put it as the second
-motivation, etc. We have ordered the factors by giving 3 points for a first
-motivation response, 2 points for a second motivation, etc. and then sorting by
+motivation, \etc We have ordered the factors by giving 3 points for a first
+motivation response, 2 points for a second motivation, \etc and then sorting by
 the sum.

### 2013-01-06 sections/introduction.tex
done with section 6

@@
 
-Nowadays, many organizations are adopting more lightweight code review
+Nowadays many organizations are adopting more lightweight code review
 practices to limit the inefficiencies of inspections. In particular, there is a
@@
 review~\cite{rigby2012open}. In the context of this paper, we define
-\emph{Modern Code Review}, as review that is (1) informal (in contrast to
-Fagan-style), (2) tool-based, and that (3) occurs regularly in practice
-nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
-Facebook~\cite{tsotsis2011online}, and in other companies and OSS projects~\cite{gerrit2012online}.
+\emph{Modern Code Review}, as review that is \begin{inparaenum}[(1)]
+\item informal (in contrast to Fagan-style), 
+\item tool-based, and that 
+\item occurs regularly in practice nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
+Facebook~\cite{tsotsis2011online}, and in other companies and OSS projects~\cite{gerrit2012online}. \end{inparaenum}
 
@@
 
-We set up our study as an explorative investigation. We started without \emph{a
+We set up our study as an exploratory investigation. We started without \emph{a
 priori} hypotheses regarding how and why code review should be performed, with
@@
 how reviews are conducted in practice, and what the actual outcomes and
-challenges are. To that end, we (1) observed 17 industrial developers
-performing code review with various degrees of experience and seniority across
-16 separate product teams with distinct reviewing cultures and policies; (2)
-interviewed these developers using a semi-structured interviews; (3) manually
-inspected and classified the content of 570 comments in discussions contained
-within code reviews; and (4) surveyed 165 managers and 873 programmers.
+challenges are. To that end, we \begin{inparaenum}[(1)] 
+\item observed 17 industrial developers performing code review with various degrees of experience and seniority across 16 separate product teams with distinct reviewing cultures and policies; 
+\item interviewed these developers using a semi-structured interviews; 
+\item manually inspected and classified the content of 570 comments in discussions contained within code reviews; and 
+\item surveyed 165 managers and 873 programmers.
+\end{inparaenum}
 

### 2013-01-06 sections/methodology.tex
done with section 6

@@
 Once reviewers open a CodeFlow review, they interact with it via a single
-desktop window (Figure~\ref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
+desktop window (\figref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
 changed in the current submission, plus a ``description.txt'' file, which
@@
    \centering
-   \includegraphics[width=0.98\columnwidth]{figs/codeflow.pdf}
-   \caption{Screenshot of CodeFlow, the dominant code review tool used by developers at Microsoft.}
+   \includegraphics[width=0.98\columnwidth]{codeflow.pdf}
+   \caption{CodeFlow, the main code review tool used by developers at Microsoft.}
    \label{fig:codeflow:screenshot}
@@
    \centering
-   \includegraphics[width=1.3\columnwidth]{figs/methodology.pdf}
+   \includegraphics[width=1.4\columnwidth]{methodology.pdf}
    \caption{The mixed approach research method applied.}
@@
 Our research method followed a mixed approach~\cite{creswell2009research}, depicted in
-Figure~\ref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
+\figref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
 analysis of previous study, (\textbf{2}) observations and interviews with developers,
@@
 
-We contacted 100 random candidates who signed-off between 50 and 250 code
+We contacted 100 randomly selected candidates who signed-off between 50 and 250 code
 reviews since the CodeFlow release and sampled across different product teams
@@
 abstraction and identify common themes. A card sort involves three phases: In
-the (1) \emph{preparation phase}, participants of the card sort are selected and the
-cards are created; in the (2) \emph{execution phase}, cards are sorted into meaningful
-groups with a descriptive title; and in the (3) \emph{analysis phase}, abstract
-hierarchies are formed to deduce general categories.
+the \begin{inparaenum}[(1)]
+\item \emph{preparation phase}, participants of the card sort are selected and the
+cards are created; in the 
+\item \emph{execution phase}, cards are sorted into meaningful
+groups with a descriptive title; and in the 
+\item \emph{analysis phase}, abstract hierarchies are formed to deduce general categories.
+\end{inparaenum}
 
@@
 review comments into categories: We randomly sampled 200 threads with at least
-two comments (e.g., Point 4 of Figure~\ref{fig:research-method}), from the entire dataset of CodeFlow
+two comments (e.g., Point 4 of \figref{fig:research-method}), from the entire dataset of CodeFlow
 reviews, which embeds data from dozens of independent software products at
@@
 to connect the related concepts and derive the main themes. For generating the
-affinity diagram, we followed the five canonical steps: we (1) recorded the
-categories on post-it-notes, (2) spread them onto a wall, (3) sorted the
-categories based on discussions, until all are sorted and all participants
-agreed, (4) named each group with a description, and (5) captured and discussed
-the themes.
+affinity diagram, we followed the five canonical steps: we \begin{inparaenum}[(1)] 
+\item recorded the categories on post-it-notes, 
+\item spread them onto a wall, 
+\item sorted the categories based on discussions, until all are sorted and all participants agreed, 
+\item named each group with a description, and 
+\item captured and discussed the themes.
+\end{inparaenum}
 
@@
 Thus, we had one sample of first level managers who participated in review,
-another sample of second level managers who participated in reviews, \etc.  The
+another sample of second level managers who participated in reviews, \etc  The
 first survey was a short survey comprising 6 questions (all optional), which we

### 2013-01-06 sections/outcomes.tex
done up to section IV (partially included)
done with section 6

--- 2013-01-05 sections/challenges.tex
@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction} \label{sec:introduction}
+\section{The Outcomes of Code Reviews} \label{sec:outcomes}
 %%%%%%%%%%%%%%%%%%%%%%
+
+Our second research question seeks to understand what the actual outcomes of code reviews are, and whether they match the motivations and expectations outlined in the previous section.
+To that end, we conducted indirect field research~\cite{lethbridge2005studying} by analyzing the content of 200 threads (corresponding to 570 comments) recorded by CodeFlow.
+\figref{XXX4} shows the categories of comments found through the card sort.
+
+\textbf{Code Improvements:} The most frequent category, with 165~(29\%) comments, is ``code improvements.'' In detail, among ``code improvements'' comments we find 58 on using better code practices, 55 on removing not necessary or unused code, and 52 on improving code readability.
+
+\textbf{Defect Finding:} Although ``defect finding'' is the top motivation and expected outcome of code review for many practitioners, the category ``defect'' is the only the fourth most frequent, out of nine items, with 78 (14\%) comments.
+Among ``defect'' comments, 65 are on logical issues (\eg a wrong expression in an if clause), 6 on high-level issues, 5 on security, and 3 on wrong exception handling.
+
+\textbf{Knowledge Transfer:} Concerning the other expected outcomes of code reviews, we did not expect to find evidence about them, because of their more ``social''—thus harder to quantify—nature.
+Nevertheless, we found some (12) comments specifically about ``knowledge transfer,'' where the reviewers were directing the code change author to external resources (\eg internal documentation or websites) for learning how to tackle some issues.
+This provides additional evidence on the importance of this aspect of reviews.
+
+\subsection{Finding defects: When expectations do not meet reality}
+
+Why do we see this significant gap in frequency between ``code improvements'' and ``defects'' comments?
+Possible reasons may be that our sample of 570 comments is too small to represent the population, that the submitted changes might require less need fixing of ``real'' defects than of small code improvements, or that programmers could consider ``code improvements'' as actual defects.
+However, by triangulating these numbers with the interview discussions, the survey answers, and the other categories of comments, another reason seems to justify this situation.
+First, we start by noting that most of the comments on ``defects'' regard uncomplicated logical errors, e.g., corner cases, common configuration values, or operator precedence.
+Then, from interview data, we see that: \begin{inparaenum}[(1)]
+\item most interviewees explained how, with tool-based code reviews, most of the found defects regard \quotation{logic issues--where the author might not have considered a particular or corner case}; 
+\item some interviewees complained that the quality of code reviews is low, because reviewers only look for easy errors: \quotation{{\normalfont [Some reviewers]} focus on formatting mistakes because they are easy {\normalfont [...]}, but it doesn’t really help.
+{\normalfont [...]} In some ways it’s kind of embarrassing if someone asks you to do a code review and all you can find are formatting mistakes when there are real mistakes to be found}; and
+\item other interviewees admitted that if the code is not among their codebase, they look at \quotation{obvious bugs (such as, exception handling).} Finally, managers mentioned \quotation{catching early obvious bugs} or \quotation{finding obvious inefficiencies or errors} as reasons for doing code review.
+\end{inparaenum}
+These points illustrate that the reason for the gap between the number of comments on ``code improvements'' and on ``defects'' is not to be found in problems in the sample or in classification misconceptions, but it is rather just additional corroborating evidence that the outcome of code review does not match the main expectation of both programmers and managers—finding defects.
+Review comments about defects are few, comprising one-eighth of the total in our sample, and mostly address ``micro'' level and superficial concerns; while programmers and managers would expect more insightful remarks on conceptual and design level issues.
+Why does this happen?
+The high frequency of understanding comments hints at the answer to our question, addressed in the next section.

### 2013-01-06 sections/implications.tex
done up to section IV (partially included)
done with section 7

@@
 %%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Recommendations for Practitioners}
+
+From our work we derive recommendations to developers:
+
+\textbf{Quality Assurance:} There is a mismatch between the expectations and the actual outcomes of code reviews.
+From our study, review does not result in identifying defects as often as project members would like and even more rarely detects deep, subtle, or ``macro'' level issues.
+Relying on code review in this way for quality assurance may be fraught.
+
+\textbf{Understanding:} When reviewers have \emph{a priori} knowledge of the context and the code, they complete reviews more quickly and provide more valuable feedback to the author.
+Teams should aim to increase the breadth of understanding of developers (if the author of a change is the only expert, she has no potential reviewers) and change authors should include code owners and others with understanding as much as possible when using review to identify defects.
+Developers indicated that when the author provided context and direction to them in a review, they could respond better and faster.
+
+\textbf{Beyond Defects:} Modern code reviews provide benefits beyond finding defects.
+Code review can be used to improve code style, find alternative solutions, increase learning, share code ownership, \etc This should guide code review policies.
+
+\textbf{Communication:} Despite the growth of tools for supporting code reviews, developers still have need of richer communication than comments annotating the changed code when reviewing.
+Teams should provide mechanisms for in-person or, at least, synchronous communication.
+
+\subsection{Implications for Researchers}
+
+Our work uncovered aspects of code review---beyond our research questions---that deserve further study:
+
+\textbf{Automate Code Review Tasks:} We observed that many code review comments were related to ``code improvement'' concerns and low-level ``micro'' defects.
+Identifying both of these are problems that research has begun to solve.
+Tools for enforcing team code conventions, checking for typos, and identifying dead code already exist.
+Even more advanced tasks such as checking boundary conditions or catching common mistakes have been shown to work in practice on real code.
+For example Google experimented with adding FindBugs to their review process, though little is reported about the results~\cite{ayewah2007using}.
+Automating these tasks frees reviewers to look for deeper, more subtle defects.
+Code review is fertile ground to have an impact with code analysis tools.
+
+\textbf{Program Comprehension in Practice:} We identified context and change understanding as challenges that developers face when reviewing, with a direct relationship to the quality of review comments.
+Interestingly, modern IDEs ship with many tools to aid context and understanding, and there is an entire conference (ICPC) devoted to code comprehension, yet all current code review tools we know of show a highlighted diff of the changed files to a reviewer with no additional tool support.
+The most common motivation that we have seen for code comprehension research is a developer that is working on new code, but we argue that reviewers reviewing code they have not seen before may be more common than a developer working on new code.
+This is a ripe opportunity for code understanding researchers to have impact on real world scenarios.
+
+\textbf{Socio-technical Effects:} Awareness and learning were cited as motivations for code review, but these outcomes are difficult to observe from traces in reviews.
+We did not investigate these further, but studies can be designed and carried out to determine if and how awareness and learning increase as a result of being involved in code review.

### 2013-01-06 sections/conclusion.tex
done up to section IV (partially included)
done with section 8 and 9

--- 2013-01-05 sections/challenges.tex
@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction} \label{sec:introduction}
+\section{Conclusion} \label{sec:conclusion}
 %%%%%%%%%%%%%%%%%%%%%%
+
+In this work, we have investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations.
+We identified understanding as a key component and provided recommendations to both practitioners and researchers.
+It is our hope that the insights we have discovered lead to more effective review in practice and improved tools, based on research, to aid developers perform code reviews.

### 2013-01-06 sections/limitations.tex
done up to section IV (partially included)
done with section 8 and 9

@@
 %%%%%%%%%%%%%%%%%%%%%%
+
+As a qualitative study, gauging the validity of our findings is a difficult undertaking~\cite{golafshani2003understanding}.
+While we have endeavored to uncover and report the expectations, outcomes, and challenges of code review, limitations may exist.
+We describe them with the steps that we took to increase confidence and validity.
+
+To achieve a comprehensive view of code review, we triangulated by collecting and comparing results from multiple sources.
+For example, we found strong agreement among the results of expectations collected from interviews, surveys of manager, and surveys of developers.
+By starting with exploratory interviews of a smaller set of subjects (17) followed by open coding to extract themes, we identified core questions that we addressed to a larger audience via survey.
+
+One common notion is that empirical research within one company or one project provides little value for the academic community, and does not contribute to scientific development.
+Historical evidence shows otherwise.
+Flyvbjerg provides several examples of individual cases that contributed to discovery in physics, economics, and social science~\cite{flyvbjerg2006five}.
+Beveridge observed for social sciences: \quotation{More discoveries have arisen from intense observation than from statistics applied to large groups} (as quoted in Kuper and Kuper~\cite{kuper1995social}, page 95).
+This should not be interpreted as a criticism of research that focuses on large samples.
+For the development of an empirical body of knowledge as championed by Basili~\cite{basili1999building}, both types of research are essential.
+To understand code review across many contexts, we observed, interviewed, surveyed, and examined code reviews from developers across a diverse group of software teams that work with codebases in various domains, of varying sizes, and with varying processes.
+
+Concerning the representativeness of our results in other contexts, other companies and OSS use tools similar to CodeFlow~\cite{gerrit2012online, tsotsis2011online, kennedy2006online}.
+However, team dynamics may differ.
+The need for code understanding may already be met in contexts where projects are smaller or there is shared code ownership and a broad system understanding across the team.
+We found that higher levels of understanding lead to more informative comments, which identify defects or aid the author in other ways so review in these contexts may uncover more defects.
+In OSS contexts, project-specific expertise often must be demonstrated prior to being accepted as a ``core committer''~\cite{bird2007open}, so learning may not be as important or frequent an outcome for review.
+
+In this work, we have used discussions within CodeFlow to identify and quantify outcomes of code review.
+However, some motivations that managers and developers described are not easily observable because they leave little trace.
+For example, determining how often code review improves team awareness or transfers knowledge is difficult to assess from the discussions in reviews.
+For these outcomes, we have responses indicating that they occur, but not ``hard evidence.''
+
+Based on review comments, survey responses, and interviews, we know that in-person discussions occurred frequently.
+While we are unable to compare frequency of these events to other outcomes as we can with events that are recorded in CodeFlow, we know that they most often occurred to address understanding needs.

### 2013-01-06 sections/challenges.tex
put emphasis instead of quotes for categories

@@
 
-Our third research question seeks to understand the main challenges faced by the reviewers when performing modern code reviews, also with respect to the expected outcomes.
-We also seek to uncover the reason behind the mismatch between expectations and actual outcomes on finding defects in code reviews.
+Our third research question seeks to understand the main challenges faced by reviewers when performing modern code reviews, also with respect to the expected outcomes.
+We also seek to uncover the reasons behind the mismatch between expectations and actual outcomes on finding defects in reviews.
 
@@
 Do understanding needs change with the expected outcome of code review?
-We included a question in the programmers’ survey to know how much understanding they needed to achieve each of the motivations listed in \figref{XXX3}.
+We included a question in the programmers' survey to know how much understanding they needed to achieve each of the motivations listed in \figref{XXX3}.
 The outcome of the question is summarized in \figref{XXX5}.
 The respondents could answer with a four values Likert's scale, by selecting the understanding of the change they felt was required to achieve the specific outcome.
-The most difficult task from the understanding perspective is ``finding defects,'' immediately followed by ``alternative solutions.'' Both clearly stand out from the other items.
-The gap in understanding needs between ``finding defects'' and ``code improvement'' seems to corroborate our hypothesis that the difference in the number of comments about these two items in review comments is mostly due to understanding issues.
-Thus, if managers and developers want code review to match their need for ``finding defects,'' context and change understanding must be improved.
+The most difficult task from the understanding perspective is \emph{finding defects}, immediately followed by \emph{alternative solutions}.
+Both clearly stand out from the other items.
+The gap in understanding needs between \emph{finding defects} and \emph{code improvement} seems to corroborate our hypothesis that the difference in the number of comments about these two items in review comments is mostly due to understanding issues.
+Thus, if managers and developers want code review to match their need for \emph{finding defects}, context and change understanding must be improved.
 
@@
 To better comprehend this situation, we included in our interview guideline a question about how the interviewees start code reviews.
-Participants explained that when they own or are very familiar with the files being changed, they have a better context and it is easier for them to understand the change submitted: \quotation{when doing code review I start with things I am familiar with, so it is easier to see what is going on.} When they are file owners, they often do not even need to read the description, but they \quotation{go directly to the files they own.} On the contrary, when they do not own files, or have to review new files, they need more information and try to get it from the description, which is deemed good when it states \quotation{what was changed and why.}
+Participants explained that when they own or are very familiar with the files being changed, they have a better context and it is easier for them to understand the change submitted: \quotation{when doing code review I start with things I am familiar with, so it is easier to see what is going on.} When they are file owners, they often do not need to read the description, but they \quotation{go directly to the files they own.} On the contrary, when they do not own files, or have to review new files, they need more information and try to get it from the description, which is deemed good when it states \quotation{what was changed and why.}
 
@@
 
-Most of the respondents (798, 91\%) answered positively to the first question, motivating it with the fact that it takes time to familiarize with the code and \quotation{learn enough about the files being modified to understand their purpose, invariants, APIs, etc.,} because \quotation{big-picture impact analysis requires contextual understanding.
+Most of the respondents (798, \ie 91\%) answered positively to the first question, motivating it with the fact that it takes time to familiarize with the code and \quotation{learn enough about the files being modified to understand their purpose, invariants, APIs, etc.,} because \quotation{big-picture impact analysis requires contextual understanding.
 When reviewing a small, unfamiliar change, it is often necessary to read through much more code than that being reviewed.} The comment of a developer anticipates the answer to the second question: \quotation{It takes a lot longer to understand unknown code, but even then understanding isn't very deep.

### 2013-01-06 sections/conclusion.tex
put emphasis instead of quotes for categories

@@
 
-In this work, we have investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations.
+In this work, we investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations.
 We identified understanding as a key component and provided recommendations to both practitioners and researchers.

### 2013-01-06 sections/expectations.tex
put emphasis instead of quotes for categories

@@
 
-Almost all the managers included ``finding defects'' as one of the reasons for
+Almost all the managers included \emph{finding defects} as one of the reasons for
 doing code reviews; for 44\% of the managers, it is the top reason. Managers
@@
 is in place}) and high level concerns (\eg \quotation{catch errors in
-design}). Concerning surveyed developers/testers, ``finding defects'' is the
+design}). Concerning surveyed developers/testers, \emph{finding defects} is the
 first motivation for code review for 383 of the programmers (44\%), second
@@
 
-Nevertheless, even though ``finding defects'' emerged from our data as a strong
+Nevertheless, even though \emph{finding defects} emerged from our data as a strong
 motivation (the first for almost half of the programmers and managers),
@@
 
-Programmers ranked ``code improvement'' as an important motivation for code
-review, close to ``finding defects:'' This is the primary motivation for 337
-programmers (39\%), the second for 208 (24\%), and the third for 135 (15\%).
-Managers reported code improvements as their primary motivation in 51 cases
-(31\%). One manager wrote how code review in her view is a
+Programmers ranked \emph{code improvement} as an important motivation for code
+review, close to \emph{finding defects}: This is the primary motivation for 337~(39\%)
+programmers, the second for 208~(24\%), and the third for 135~(15\%).
+Managers reported \emph{code improvement} as their primary motivation in 51~(31\%) cases. One manager wrote how code review in her view is a
 \quotation{discipline of explaining your code to your peers {\normalfont[that]} drives a
@@
 the team conventions, for example in terms of code formatting and in terms of
-function and variable naming. Some programmers use the ``code improvement''
+function and variable naming. Some programmers use a ``code improvement''
 check as a first step when doing code review: \quotation{the first basic pass
@@
 The interviews also gave us a glimpse of the connection between the quality of
-code reviews and ``code improvement'' comments. Such comments seem easier to
+code reviews and code improvement comments. Such comments seem easier to
 write and sometimes interviewees mentioned them as the way reviewers use to
@@
 
-``Alternative solutions'' regard changes and comments on improving the
+\emph{Alternative solutions} regard changes and comments on improving the
 submitted code by adopting an idea that leads to a better implementation. This
@@
 
-All the interviewees but one motivated their code reviews also from a learning,
-or ``knowledge transfer,'' perspective. With the words of a senior developer:
+All the interviewees but one motivated their code reviews also from a learning---or \emph{knowledge transfer}---perspective. With the words of a senior developer:
 \quotation{one of the things that should be happening with code reviews over
@@
 that was not as good code review as it could have been.} Although we did not
-include questions related to ``knowledge transfer'' in our interview guideline,
+include questions related to \emph{knowledge transfer} in our interview guideline,
 this topic kept emerging spontaneously from each meeting, thus underscoring its
@@
 
-Managers included ``knowledge transfer'' as one of the reasons for code review,
+Managers included \emph{knowledge transfer} as one of the reasons for code review,
 although never as the top motivation. They mostly wrote about code review as an
@@
 
-Programmers answering the survey declared ``knowledge transfer'' to be their
+Programmers answering the survey declared \emph{knowledge transfer} to be their
 first motivation for code review in 73~(8\%) cases, their second in 119~(14\%),
@@
 
-The 873 programmers answering the survey ranked ``team awareness and
-transparency'' very close to ``knowledge transfer.'' In fact, the two concepts
+The 873 programmers answering the survey ranked \emph{team awareness and
+transparency} very close to \emph{knowledge transfer}. In fact, the two concepts
 appeared logically related also in the interviews; for example one tester,
@@
 
-Although team awareness and transparency emerged from our data as clearly
+Although \emph{team awareness and transparency} emerged from our data as clearly
 promoted by the code review process, academic research seems to have given
@@
 
-The concept of ``shared code ownership'' is closely related to ``team awareness
-and transparency,'' but it has a stronger connotation toward active
+The concept of \emph{shared code ownership} is closely related to \emph{team awareness
+and transparency}, but it has a stronger connotation toward active
 collaboration and overlapping coding activities. Programmers and managers
@@
 
-In the programmers' survey, 51~(6\%) respondents marked ``share code
-ownership'' as their first motivation, 100~(11\%) as their second, and 91~(10\%)
+In the programmers' survey, 51~(6\%) respondents marked \emph{share code
+ownership} as their first motivation, 100~(11\%) as their second, and 91~(10\%)
 as their third.
@@
 
-We discussed the five most prominent motivations, which show that ``finding
-defects'' is the top motivation, although participants believe that code review
+We discussed the five most prominent motivations, which show that \emph{finding
+defects} is the top motivation, although participants believe that code review
 brings other benefits. The first two motivations were already popular in
@@
 inspections; on the contrary, the other motivations are still unexplored,
-especially those regarding more ``social'' benefits on the team, such as shared
+especially those regarding more social benefits on the team, such as shared
 code ownership.

### 2013-01-06 sections/implications.tex
put emphasis instead of quotes for categories

@@
 
-\textbf{Automate Code Review Tasks:} We observed that many code review comments were related to ``code improvement'' concerns and low-level ``micro'' defects.
+\textbf{Automate Code Review Tasks:} We observed that many code review comments were related to \emph{code improvement} concerns and low-level ``micro'' defects.
 Identifying both of these are problems that research has begun to solve.

### 2013-01-06 sections/limitations.tex
put emphasis instead of quotes for categories

@@
 Based on review comments, survey responses, and interviews, we know that in-person discussions occurred frequently.
-While we are unable to compare frequency of these events to other outcomes as we can with events that are recorded in CodeFlow, we know that they most often occurred to address understanding needs.
+While we cannot compare frequency of these events to other outcomes as we can with events recorded in CodeFlow, we know that they most often occurred to address understanding needs.

### 2013-01-06 sections/methodology.tex
put emphasis instead of quotes for categories

@@
 \item sorted the categories based on discussions, until all are sorted and all participants agreed, 
-\item named each group with a description, and 
+\item named each group, and 
 \item captured and discussed the themes.
@@
 first survey was a short survey comprising 6 questions (all optional), which we
-sent to 600 managers that had at least ten direct or indirect reporting
-developers who used CodeFlow in the past. The central focus was the open
+sent to 600 managers that had at least 10 direct or indirect reporting
+developers who used CodeFlow. The central focus was the open
 question asking to enumerate the main motivations for doing code reviews in

### 2013-01-06 sections/outcomes.tex
put emphasis instead of quotes for categories

--- 2013-01-05 sections/challenges.tex
@@
-%!TEX root = paper.tex
+%!TEX root = ../paper.tex
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction} \label{sec:introduction}
+\section{The Outcomes of Code Reviews} \label{sec:outcomes}
 %%%%%%%%%%%%%%%%%%%%%%
+
+Our second research question seeks to understand what the actual outcomes of code reviews are, and whether they match the motivations and expectations outlined in the previous section.
+To that end, we conducted indirect field research~\cite{lethbridge2005studying} by analyzing the content of 200 threads (corresponding to 570 comments) recorded by CodeFlow.
+\figref{XXX4} shows the categories of comments found through the card sort.
+
+\textbf{Code Improvements:} The most frequent category, with 165~(29\%) comments, is \emph{code improvements}.
+In detail, among \emph{code improvements} comments we find 58 on using better code practices, 55 on removing not necessary or unused code, and 52 on improving code readability.
+
+\textbf{Defect Finding:} Although \emph{defect finding} is the top motivation and expected outcome of code review for many practitioners, the category \emph{defect} is the only the fourth most frequent, out of nine items, with 78 (14\%) comments.
+Among \emph{defect} comments, 65 are on logical issues (\eg a wrong expression in an if clause), 6 on high-level issues, 5 on security, and 3 on wrong exception handling.
+
+\textbf{Knowledge Transfer:} Concerning the other expected outcomes of code reviews, we did not expect to find evidence about them, because of their more ``social''--thus harder to quantify—nature.
+Nevertheless, we found some (12) comments specifically about \emph{knowledge transfer}, where the reviewers were directing the code change author to external resources (\eg internal documentation or websites) for learning how to tackle some issues.
+This provides additional evidence on the importance of this aspect of reviews.
+
+\subsection{Finding defects: When expectations do not meet reality}
+
+Why do we see this significant gap in frequency between \emph{code improvements} and \emph{defects} comments?
+Possible reasons may be that our sample of 570 comments is too small to represent the population, that the submitted changes might require less need fixing of ``real'' defects than of small code improvements, or that programmers could consider \emph{code improvements} as actual defects.
+However, by triangulating these numbers with the interview discussions, the survey answers, and the other categories of comments, another reason seems to justify this situation.
+First, we start by noting that most of the comments on \emph{defects} regard uncomplicated logical errors, \eg corner cases, common configuration values, or operator precedence.
+Then, from interview data, we see that: \begin{inparaenum}[(1)]
+\item most interviewees explained how, with tool-based code reviews, most of the found defects regard \quotation{logic issues--where the author might not have considered a particular or corner case}; 
+\item some interviewees complained that the quality of code reviews is low, because reviewers only look for easy errors: \quotation{{\normalfont [Some reviewers]} focus on formatting mistakes because they are easy {\normalfont [...]}, but it doesn't really help.
+{\normalfont [...]} In some ways it's kind of embarrassing if someone asks you to do a code review and all you can find are formatting mistakes when there are real mistakes to be found}; and
+\item other interviewees admitted that if the code is not among their codebase, they look at \quotation{obvious bugs (such as, exception handling).} Finally, managers mentioned \quotation{catching early obvious bugs} or \quotation{finding obvious inefficiencies or errors} as reasons for doing code review.
+\end{inparaenum}
+These points illustrate that the reason for the gap between the number of comments on \emph{code improvements} and on \emph{defects} is not to be found in problems in the sample or in classification misconceptions, but it is rather just additional corroborating evidence that the outcome of code review does not match the main expectation of both programmers and managers—finding defects.
+Review comments about defects are few, comprising one-eighth of the total in our sample, and mostly address ``micro'' level and superficial concerns; while programmers and managers would expect more insightful remarks on conceptual and design level issues.
+Why does this happen?
+The high frequency of understanding comments hints at the answer to our question, addressed in the next section.

### 2013-01-08 sections/challenges.tex
Added figures and references to them

@@
 
+\begin{figure}[t] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=\columnwidth]{understanding.pdf}
+   \vspace{-1.5em}
+   \caption{Developers’ responses in surveys of the amount of code understanding for code review outcomes.}
+   \label{fig:understanding}
+   \vspace{-1.5em}
+\end{figure}
+
 From interviews, no other code review challenge emerged as clearly as understanding the submitted change.
-Even though scheduling and time issues also appeared challenging, we could always trace them back to the first challenge—through the words of a tester: \quotation{understanding the code takes most of the reviewing time.} On the same note, in the code review comments we analyzed, the second most frequent category concerns understanding.
+Even though scheduling and time issues also appeared challenging, we could always trace them back to the first challenge through the words of a tester: \quotation{understanding the code takes most of the reviewing time.} On the same note, in the code review comments we analyzed, the second most frequent category concerns understanding.
 This category includes clarification questions and doubts raised by the reviewers who want to grasp the rationale of the changes done on the code, and the corresponding clarification answers.
@@
 Do understanding needs change with the expected outcome of code review?
-We included a question in the programmers' survey to know how much understanding they needed to achieve each of the motivations listed in \figref{XXX3}.
-The outcome of the question is summarized in \figref{XXX5}.
+We included a question in the programmers' survey to know how much understanding they needed to achieve each of the motivations listed in \figref{fig:dev-motivations}.
+The outcome of the question is summarized in \figref{fig:understanding}.
 The respondents could answer with a four values Likert's scale, by selecting the understanding of the change they felt was required to achieve the specific outcome.

### 2013-01-08 scripts/dev_motivations.py
added data files and code for generating figures

@@ -1,0 +1,29 @@
+import re
+
+p = re.compile(r"/|\(")
+
+buckets = {}
+
+for line in open("survey_developers.tsv").readlines()[1:]:
+	parts=line.split("\t")
+	field = parts[8]
+	categories = field.split(";#")
+	for category in categories:
+		category = category.replace('"', '')
+		name = p.split(category)[0]
+		print name
+		if not name in buckets.keys():
+			buckets[name] = [0, 0, 0]
+		if "First" in category:
+			buckets[name][0] += 1;
+		if "Second" in category:
+			buckets[name][1] += 1;
+		if "Third" in category:
+			buckets[name][2] += 1;
+
+fd = open("dev_motivations.tsv", "w")
+for k, v in buckets.items():
+	print >> fd, k + "\t" + str(v[0]) + "\t" + str(v[1]) + "\t" + str(v[2])
+	print k + "\t" + str(v[0]) + "\t" + str(v[1]) + "\t" + str(v[2])
+
+

### 2013-01-08 scripts/dev_motivations_plot.r

@@ -1,0 +1,124 @@
+
+
+motivations = function()
+{
+setwd("c:/Users/cbird/Documents/Projects/codereview/")
+a = read.table("dev_motivations.tsv", sep="\t", header=1)
+pdf("C:/Users/cbird/Documents/GitHub/code_review_icse13/figs/dev_motivations.pdf", width=8, height=6)
+old = par("mar")
+par(mar=c(5, 13, 4, 2))
+cols = gray.colors(3)
+k=barplot(t(as.matrix(a[10:1,2:4])), 
+        horiz=T, names.arg=a[10:1,1], xlim=c(0,700),
+        main = "Ranked Motivations From Developers",
+        col=cols,
+        las=2, args.legend=list(horiz=TRUE, bty="n", x=650, y=13.3, cex=1.25),
+          legend.text=c("Top", "Second", "Third"), cex.names=1.25, cex.main=1.5)
+#legend(x=350, y=13, legend=c("Top", "Second", "Third"), horiz=TRUE, fill=cols)
+dev.off();
+par(mar=old)
+}
+motivations = function(dopdf=FALSE)
+{
+  setwd("c:/Users/cbird/Documents/Projects/codereview/")
+  a = read.table("dev_motivations.tsv", sep="\t", header=1)
+  if (dopdf)
+  {
+    pdf("C:/Users/cbird/Documents/GitHub/code_review_icse13/figs/dev_motivations.pdf", width=9, height=6)
+  }
+  #old = par("mar")
+  #par(mar=c(5, 13, 4, 2))
+  cols = gray.colors(3)
+  f = as.matrix(a[10:1,c(2,3,4)])
+  rownames(f) = a$category[10:1]
+  plot(barchart(f, col=cols[c(3,2,1)],
+                xlab=list("Responses", cex=1.3),
+              
+                
+                main=list("Ranked Motivations From Developers", cex=1.3),
+                scales=list(cex=1.3), 
+                            #y=list(cex=1.3)),
+                key=list(
+                  text=list(c("Top", "Second", "Third"), cex=1.3), 
+                  rectangles=list(col=cols[c(3, 2, 1)]),
+                  columns=3)
+  )
+  )
+  if (dopdf)
+  {
+    dev.off()
+  }
+  #par(mar=old)
+}
+
+
+
+understanding = function(dopdf=FALSE)
+{
+  setwd("c:/Users/cbird/Documents/Projects/codereview/")
+  a = read.table("understanding.tsv", sep="\t", header=1)
[…]

### 2013-01-08 scripts/understanding.py

@@ -1,0 +1,33 @@
+import re
+
+p = re.compile(r"/|\(")
+
+buckets = {}
+
+for line in open("survey_developers.tsv").readlines()[1:]:
+	parts=line.split("\t")
+	field = parts[17]
+	categories = field.split(";#")
+	for category in categories:
+		category = category.replace('"', '')
+		name = p.split(category)[0].strip()
+		if not name:
+			continue
+		#print name
+		if not name in buckets.keys():
+			buckets[name] = [0, 0, 0, 0]
+		if "No understanding" in category:
+			buckets[name][0] += 1;
+		if "Little understanding" in category:
+			buckets[name][1] += 1;
+		if "High understanding" in category:
+			buckets[name][2] += 1;
+		if "Complete understanding" in category:
+			buckets[name][3] += 1;
+
+fd = open("understanding.tsv", "w")
+for k, v in buckets.items():
+	print >> fd, k + "\t" + str(v[0]) + "\t" + str(v[1]) + "\t" + str(v[2]) + "\t" + str(v[3])
+	print k + "\t" + str(v[0]) + "\t" + str(v[1]) + "\t" + str(v[2]) + "\t" + str(v[3])
+
+

### 2013-01-10 reviews.tex
added reviews

\documentclass[a4paper,10pt]{scrartcl}

\usepackage{times,graphicx,alltt,enumerate,amsmath,amssymb,eurosym,hyperref,a4wide,cite,xspace,ifthen,url,hyperref,color,soul,pifont}

\newcounter{revnumber}

\newenvironment{review}[1]{
	\addtocounter{revnumber}{1}
	\subsection*{Review \arabic{revnumber} - #1
	\makeatother}
	}

\newenvironment{myindentpar}[1]%
{\begin{list}{}%
         {\setlength{\leftmargin}{#1}}%
         \item[]%
}
{\end{list}}

\newcommand\serieheader[6]{
	{\noindent \bf \Large \sffamily {#1} {#2}}\\[2mm] {\sf #3} \hfill {\sf #5}\\[2mm] {\sf #4} \hfill {\sf #6} \noindent \rule{\linewidth}{0.3mm}
	}

\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
{\newcommand{\nb}[2]{\fbox{\bfseries\sffamily\scriptsize#1}{\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}}}
{\newcommand{\nb}[2]{}}
\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\newcommand\CB[1]{\nb{Chris}{#1}}
\newcommand\cb[1]{\nb{Chris}{#1}}
\newcommand\ab[1]{\nb{Alberto}{#1}}
\newcommand\AB[1]{\nb{Alberto}{#1}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{review} {}

\noindent{\bf Summary}

The authors presents the results of an empirical study at Microsoft that
investigated code review practices. The authors use the term ``modern
code review'' meaning that such reviews are performed informally, are
supported with tools, and are performed frequently in everyday software
engineering practice. The authors observed 17 developers in 16 product
teams and analyzed 570 review comments to derive recommendations for
practitioners and researchers working of review processes and tools.
They study the motivations of developers, actual outcomes of reviews,
and challenges experienced. The study shows that besides defect
detection knowledge transfer, team awareness, and improved solutions are
the key factors that motivate developers to perform code reviews.\\

\noindent{\bf Evaluation}

The paper addresses a (still) interesting topic. Structured inspection
techniques proposed in the 1970s are rarely performed in software
development in their original form today. Instead companies are nowadays
using more lightweight techniques to perform code reviews. Analyzing
current review practices and analyzing the benefits and actual use in
current real-world development processes is an important and interesting
topic of research.

The paper is well written and motivated and I liked reading it. I liked
particularly that the authors takes a very practical and realistic view
on code reviews. The research method involving the review of a previous
study, observations, interviews, card sorting techniques, affinity
diagrams and survey for validation is well described and reasonable.

The results rather confirm what can be expected but are not really
surprising in my view. It is known already that code reviewing helps
finding defects, maintaining team awareness, improving code quality, and
transferring knowledge among team members. The recommendations and
implications summarize the detailed results but are also rather
high-level. While interesting I doubt that a lot can be learned.

Regarding threats to validity I agree with the authors that empirical
research done in one company can contribute to scientific development if
it is based on careful observation. However, the study might be biased
by the CodeFlow code review tool that was used by developers participating
in the study. The authors briefly discuss this issue but this discussion
should be extended in my view. Furthermore, only 17 users participated
in the study. The authors write that they sampled across different
product teams to address different perspectives and to cover different
levels of experience. Still 17 participants is not a high number given that
the CodeFlow user base is 28.000 according to the authors.

The authors investigate ad-hoc review processes. However, I missed a
discussion in the paper of how the results relate to more conventional
techniques traditionally used in code reviews. For instance, would
checklists or reading techniques be useful in the described development
context?

The paper lists a number of existing studies and papers on software
inspection and code reviews but also misses important papers from the
literature. The following references should be considered:

\begin{enumerate}

\item Parnas \& Weiss, D.M. Active Design Reviews: Principles and Practice. JSS
7, 259-265 (1987)

\item Perpich et al., Anywhere, Anytime Code Inspections: Using the Web to
Remove Inspection Bottlenecks in Large-Scale Software Development. ICSE
1997

\item Biffl el al., A family of experiments to investigate the effects of
groupware for software inspection. ASE Journal 13(3): 373-394 (2006)

\item Johnson \& Tjahjono, Assessing software review meetings: A controlled
experimental study using CSRS. ICSE 1997

\item Lanubile \& Mallardo, An Empirical Study of Web-Based Inspection
Meetings. ACM/IEEE Int. Symp. on Empirical Software Engineering, Rome,
pp. 244-251 (2003).

\end{enumerate}

\end{review}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{review}{}

\noindent{\bf Summary}

The authors examine the expectations, outcomes and challenges of code review as
applied at Microsoft using the CodeFlow tool by surveying and interviewing a
large collection of developers and managers. The study had several components:
analyzing the results of a prior study, observing developers and interviewing
developers, performing a card sort of interview data, examining and sorting
review comments, creating an affinity diagram, and surveying managers and
developers. The authors examine 3 research questions and then provide guidance
on what areas should be further researched.\\

\noindent{\bf Evaluation}

Positives:

\begin{itemize}

\item[+] The authors undertook a large study using multiple techniques and sources of
data, analyzing it using widely applied techniques, and then looking for trends

\item[+] The authors identify differences in the expectations and outcomes of managers
and developers

\item[+] The authors place their work in the context of related work

\item[+] The authors identify the work that should be approached in the future

\item[+] The paper is well written and the approach is explained well

\item[+] The take away message is clear - though developers and managers plan to use
code reviews to find faults, that is not really the main thing that happens.
Reviews are mainly used for code improvement and understanding, with
understanding being a major issue for developers (indicating a need for further
research).

\end{itemize}

Opportunities for improvement:

\begin{itemize}

\item[p.2] Please describe the initial in-field observations

\item[p.3] was there a Hawthorne effect to the ``thought aloud'' portion of the work?

\item[p.4] top left column - please show a sample card?

\item[p.4] top left - is there not possible bias from the first author sorting the
cards to form initial themes?

\item[p.4] middle of left column - how time consuming were the card sorts?

\end{itemize}

Of lesser importance:

There are some typos and minor grammatical issues - a thorough proofread of the
paper will remedy this

p. 9 need text under section VII header

\end{review}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{review}{}

\noindent{\bf Summary}

This paper reports the results of an empirical study of what the authors call
``modern code inspections,'' i.e., lightweight, asynchronous inspections carried
out by individuals with the aid of online tools. The authors used a
multi-method approach, employing interviews, a survey, and analyzing review
comments captured by the tool. The results describe the purposes for which
inspections are carried out, what actually happens in the inspection process as
revealed by the comments, and a broad sampling of managers and developers via
the survey. Among the results are that finding defects is (unsurprisingly) the
most frequently mentioned purpose, but unexpectedly, several other purposes are
very nearly as common, such as code improvement, finding alternative solutions,
and learning about the code base. In fact, inspections appear not to be
terribly effective in finding any but the shallowest and most obvious bugs,
unless the code inspectors are already familiar with the code they are reading.
The authors provide recommendations for tool improvements and for future
research that address these overlooked, and potentially more important,
purposes.\\

\noindent{\bf Evaluation}

This is an excellent paper. The topic of inspections, one might think at first,
has been the subject of so much empirical study that there is little left to
learn. This paper shows definitively that there is much more to learn. The
methods were very well suited to the exploratory nature of the work, and the
use of multiple methods, each with different strengths and weaknesses, provides
a very convincing base of evidence for the claims in the paper. For example,
developers pursue a number of purposes in participating in inspections, as the
interviews clearly bring out. The survey of managers provides a contrasting
view, as they share some of these purposes (e.g., finding bugs) but not others
(e.g., finding alternative solutions). The analysis of comments enriches the
results further, showing what developers communicated about, and the sorts of
issues that appeared to absorb their time and energy (e.g., code
understanding), and those that did not (e.g., trying to find deep bugs).

The paper makes great use of triangulation, i.e., seeking additional evidence
to support or refute a tentative interpretation. One of the most interesting
was the observation, from the interviews and comments, that usually only
superficial bugs were found, and another observation that unfamiliar code was
very hard to understand. In the survey of developers, they asked for each of
the inspection outcomes (e.g., finding bugs, learning about the code, improving
the code) how much code understanding was required (rated on a scale from
"none" to "complete"). Interestingly, "finding defects" came out as requiring
the greatest level of understanding, and along with "alternative solutions"
were so far above the others, that they seem qualitatively very different.
Other outcomes one might expect to be similar, such as "code improvement" were
far lower. This finding not only reinforces the earlier results about finding
only superficial bugs and understanding being the most difficult part of the
inspection, it suggests an interesting relationship between the two.

This is an important paper. It provides convincing evidence that inspections
are important, but not for the reasons we thought. In fact, under many
conditions, they are not terribly useful for finding bugs, but they accomplish
many other important purposes. As the authors point out, tools and processes
are optimized around those things that inspections are not very good at, and
they neglect important things like providing assistance for understanding the
code, creating a sense of shared ownership, and transferring knowledge. I think
this paper will change how the field thinks about inspections and what we need
to support them. In fact, it has implications beyond inspections, for example,
the about the potential value of "transparent" environments, such as GitHub,
where all coding is done out in the open, and for pair programming that in
practice can be very close to lightweight inspections, and serves many of the
same purposes.

As an aside, the paper also introduces some interesting analysis techniques.
Open sorting is an interesting variation on the standard qualitative technique
of open coding, and the way they used the affinity diagram - as a technique to
find order in the categories they derived from open coding - is an interesting
way to carry out axial coding. It might be helpful to readers to connect these
techniques with the classical qualitative methods works such as:

Strauss, A., \& Corbin, J. (1990). Basics of Qualitative Research: Grounded
Theory Procedures and Techniques. Newbury Park, CA: Sage Publications, Inc.

A few nits:

\begin{itemize}

\item[p.1] explorative --$>$ exploratory

\item[p.3]: we contacted 100 random candidates - I think you mean they were selected
randomly. The candidates might bristle at being called random . . .

\item[p.9]: ``one common notion . . . '' I don't think this notion is at all common. I
think it is quite out of step with what people believe and the kind of work
that does well in the review process. What is the source of this? I see no
citations. The sentences that follow sound needlessly defensive to me.

\end{itemize}

\end{review}

\end{document}

### 2013-01-10 reviews.tex
revised review 1

@@
 
-The results rather confirm what can be expected but are not really
-surprising in my view. It is known already that code reviewing helps
+\hl{The results} rather confirm what can be expected but \hl{are not really
+surprising in my view}. It is known already that code reviewing helps
 finding defects, maintaining team awareness, improving code quality, and
-transferring knowledge among team members. The recommendations and
+transferring knowledge among team members. \hl{The recommendations and
 implications summarize the detailed results but are also rather
-high-level. While interesting I doubt that a lot can be learned.
+high-level}. While interesting I doubt that a lot can be learned.
 
@@
 research done in one company can contribute to scientific development if
-it is based on careful observation. However, the study might be biased
-by the CodeFlow code review tool that was used by developers participating
+it is based on careful observation. However, \hl{the study might be biased
+by the CodeFlow code review tool} that was used by developers participating
 in the study. The authors briefly discuss this issue but this discussion
@@
 product teams to address different perspectives and to cover different
-levels of experience. Still 17 participants is not a high number given that
+levels of experience. Still \hl{17 participants is not a high number} given that
 the CodeFlow user base is 28.000 according to the authors.
 
-The authors investigate ad-hoc review processes. However, I missed a
+The authors investigate ad-hoc review processes. However, \hl{I missed a
 discussion in the paper of how the results relate to more conventional
-techniques traditionally used in code reviews. For instance, would
+techniques traditionally used in code reviews}. For instance, would
 checklists or reading techniques be useful in the described development
@@
 The paper lists a number of existing studies and papers on software
-inspection and code reviews but also misses important papers from the
-literature. The following references should be considered:
+inspection and code reviews but also \hl{misses important papers from the
+literature}. The following references should be considered:
 

### 2013-01-10 reviews.tex
revised review 2 and 3

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{review} {}
+\begin{review} {The Bad}
 
@@
 
-\begin{review}{}
+\begin{review}{The Ugly}
 
@@
 
-\item[p.2] Please describe the initial in-field observations
+\item[p.2] Please \hl{describe the initial in-field observations}.
 
-\item[p.3] was there a Hawthorne effect to the ``thought aloud'' portion of the work?
+\item[p.3] was there a \hl{Hawthorne effect to the ``thought aloud'' portion} of the work?
 
-\item[p.4] top left column - please show a sample card?
+\item[p.4] top left column - please \hl{show a sample card}?
 
-\item[p.4] top left - is there not possible bias from the first author sorting the
-cards to form initial themes?
+\item[p.4] top left - is there not possible \hl{bias from the first author sorting the
+cards} to form initial themes?
 
-\item[p.4] middle of left column - how time consuming were the card sorts?
+\item[p.4] middle of left column - \hl{how time consuming were the card sorts}?
 
@@
 
-There are some typos and minor grammatical issues - a thorough proofread of the
+\begin{itemize}
+
+\item[pp.] There are some typos and minor grammatical issues - a thorough proofread of the
 paper will remedy this
 
-p. 9 need text under section VII header
+\item[p.9] need text under section VII header
+
+\end{itemize}
 
@@
 
-\begin{review}{}
+\begin{review}{The Good}
 
@@
 the code) how much code understanding was required (rated on a scale from
-"none" to "complete"). Interestingly, "finding defects" came out as requiring
-the greatest level of understanding, and along with "alternative solutions"
+``none'' to ``complete''). Interestingly, ``finding defects'' came out as requiring
+the greatest level of understanding, and along with ``alternative solutions''
 were so far above the others, that they seem qualitatively very different.
-Other outcomes one might expect to be similar, such as "code improvement" were
+Other outcomes one might expect to be similar, such as ``code improvement'' were
 far lower. This finding not only reinforces the earlier results about finding
@@
 to support them. In fact, it has implications beyond inspections, for example,
-the about the potential value of "transparent" environments, such as GitHub,
+the about the potential value of ``transparent'' environments, such as GitHub,
 where all coding is done out in the open, and for pair programming that in
@@
 find order in the categories they derived from open coding - is an interesting
-way to carry out axial coding. It might be helpful to readers to connect these
-techniques with the classical qualitative methods works such as:
+way to carry out axial coding. \hl{It might be helpful to readers to connect these
+techniques with the classical qualitative methods works} such as:
 
@@
 
-\item[p.1] explorative --$>$ exploratory
-
-\item[p.3]: we contacted 100 random candidates - I think you mean they were selected
-randomly. The candidates might bristle at being called random . . .
-
 \item[p.9]: ``one common notion . . . '' I don't think this notion is at all common. I
@@
 citations. The sentences that follow sound needlessly defensive to me.
+
+\AB{I am not mentioning the two ``nits'' already fixed}
 

### 2013-01-10 sections/limitations.tex
fixed first problem

@@
 
-One common notion is that empirical research within one company or one project provides little value for the academic community, and does not contribute to scientific development.
+One potential criticism is that empirical research within one company or one project provides little value for the academic community, and does not contribute to scientific development.
 Historical evidence shows otherwise.

### 2013-03-01 sections/methodology.tex
put reference for saturation, and mentioned how we tried to minimize the Hawthorne effect for observation

@@
 performing the code review that they had been assigned. To minimize
-invasiveness we used only one observer and to encourage the participant to
-narrate their work, we asked the participants to think of us as a newcomer to
+invasiveness and the Hawthorne effect, we used only one observer, and to encourage the participant to
+narrate their work, we asked the participants to consider him as a newcomer to
 the team. In this way, most developers thought aloud without need of prompting.
@@
 Since we, as observers, have backgrounds in software development and practices
-at Microsoft, we were able to understand most of the work and where and how
+at Microsoft, we could understand most of the work and where and how
 information was obtained without inquiry.
@@
 
-Observations also reached a \emph{saturation} point, thus providing insights very similar to the earlier ones.
-For this, after the first 5-6 observations, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.
+After the first 5-6 meetings, the observations reached a \emph{saturation} point~\cite{Glas1998a}: They were providing insights very similar to the earlier ones.
+For this, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.
 

### 2013-03-01 sections/methodology.tex
talked about how we reduced bias from the first author sorting the cards
added that we do both qualitative and quantitative research

@@
 
-Our research method followed a mixed approach~\cite{creswell2009research}, depicted in
+Our research method followed a mixed qualitative and quantitative approach~\cite{creswell2009research}, depicted in
 \figref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
@@
 the cards were sorted by the first author several times to identify initial
-themes. Next, all researchers reviewed and agreed on the final set of
+themes. To reduce bias from the first author sorting the cards to form initial themes, all researchers reviewed and agreed on the final set of
 categories.

### 2013-03-01 sections/introduction.tex
removed sentence about a thing we do not actually do

@@
 \item occurs regularly in practice nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
-Facebook~\cite{tsotsis2011online}, and in other companies and OSS projects~\cite{gerrit2012online}. \end{inparaenum}
+Facebook~\cite{tsotsis2011online}, and in other organizations and OSS projects~\cite{gerrit2012online}. \end{inparaenum}
 
-This trend raises questions, such as: Can we apply the lessons learned from
-previous research on code inspections to modern code reviews? What are the
+This trend raises questions, such as: What are the
 expectations for code review nowadays? What are the actual outcomes of code

### 2013-03-01 sections/conclusion.tex
did another full pass on the paper

@@
 
-In this work, we investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations.
+We investigated modern, tool-based code review, uncovering both a wide range of motivations for review and that the outcomes do not always match those motivations.
 We identified understanding as a key component and provided recommendations to both practitioners and researchers.

### 2013-03-01 sections/introduction.tex
did another full pass on the paper

@@
 \item occurs regularly in practice nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
-Facebook~\cite{tsotsis2011online}, and in other organizations and OSS projects~\cite{gerrit2012online}. \end{inparaenum}
+Facebook~\cite{tsotsis2011online}, and in other organizations and open source software (OSS) projects~\cite{gerrit2012online}. \end{inparaenum}
 
@@
 about when to use code review and how it should fit into their development
-process. Researchers can focus their attention on practitioners' challenges to
-make code review more effective.
+process. Researchers can focus their attention on the challenges faced by practitioners to make code review more effective.
 
@@
 Our results show that, although the top motivation driving code reviews is
-still finding defects, the practice and the actual outcomes are less about
+finding defects, the practice and the actual outcomes are less about
 finding errors than expected: Defect related comments comprise a small
-proportion and mainly cover small logical low-level issues. At the same time,
+proportion and mainly cover small logical low-level issues. On the other hand,
 code review additionally provides a wide spectrum of benefits to software

### 2013-03-01 sections/methodology.tex
added reference to tech report for seeing example card
did another full pass on the paper

@@
 Our investigation of code review revolves around the following research
-questions, which we iteratively refined during our initial in-field
-observations and interviews:
+questions (iteratively refined during our initial in-field
+observations and interviews):
 
@@
 
-Once reviewers open a CodeFlow review, they interact with it via a single
+Once reviewers open a CodeFlow review, they interact with it through a single
 desktop window (\figref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
-changed in the current submission, plus a ``description.txt'' file, which
+changed in the current submission, plus a `description.txt' file, which
 contains a textual explanation of the change, written by the author. On bottom
@@
 
-CodeFlow centralizes and records all the information on code reviews on a
+CodeFlow records all the information on code reviews on a
 central server. This provides an additional data source that we used to analyze
@@
 
-Our research method followed a mixed qualitative and quantitative approach~\cite{creswell2009research}, depicted in
-\figref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
+Our research method followed a mixed qualitative and quantitative approach~\cite{creswell2009research} 
+(depicted in \figref{fig:research-method}), which collects data from different sources for triangulation: (\textbf{1})
 analysis of previous study, (\textbf{2}) observations and interviews with developers,
@@
 
-Most of the interview questions revolved around topics that are very specific
-to tool usage, and were only tangentially related to this work. We found one
+Most of the interview questions revolved around topics that were very specific
+to usage of the tool, and were only tangentially related to this work. We found one
 relevant as a starting point for our study: ``\emph{What do you hope to accomplish
@@
 
-\textbf{2. Observations and interviews with developers:} Next, we conducted a series of
+\textbf{2. Observations and interviews with developers:} Subsequently, we conducted a series of
 one-to-one meetings with developers who use CodeFlow, each taking 40-60
@@
 
-The second part of the meeting was a \emph{semi-structured} interview~\cite{taylor2010qualitative}. Semi-structured interviews make use of an \emph{interview guide} that
+The second part of the meeting was a \emph{semi-structured} interview~\cite{taylor2010qualitative}. This form of interviews makes use of an \emph{interview guide} that
 contains general groupings of topics and questions rather than a pre-determined
 exact set and order of questions.  They are often used in an exploratory
-context to ``find out what is happening [and] to seek new insights''~\cite{weiss1995learning}. 
+context to \quotation{find out what is happening \emph{[and]} to seek new insights}~\cite{weiss1995learning}. 
 The guideline was iteratively refined after each interview, in
@@
 After the first 5-6 meetings, the observations reached a \emph{saturation} point~\cite{Glas1998a}: They were providing insights very similar to the earlier ones.
-For this, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.
+For this, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a hook to talk about topics in our guideline.
 
@@
 The first author of this paper created all of the cards, from the 1,047
-coherent units in the interviews. Throughout our further analysis other
+coherent units in the interviews (an example card is shown in a technical report~\cite{bacchelli2012appendix}). Throughout our further analysis other
 researchers (the second author and external people) were involved in developing
@@
 review comments into categories: We randomly sampled 200 threads with at least
-two comments (e.g., Point 4 of \figref{fig:research-method}), from the entire dataset of CodeFlow
+two comments (\eg Point 4 of \figref{fig:research-method}), from the entire dataset of CodeFlow
 reviews, which embeds data from dozens of independent software products at
@@
 two surveys to reach a significant number of participants and to challenge our
-conclusions (The full surveys are available as a technical report~\cite{bacchelli2012appendix}). 
+conclusions (the full surveys are available as a technical report~\cite{bacchelli2012appendix}). 
 For the design of the surveys, we followed Kitchenham and

### 2013-03-01 sections/related_work.tex
did another full pass on the paper

@@
 
-Previous studies exist that have examined the practices of code inspection and
+Previous studies have examined the practices of code inspection and
 code review.  Stein \etal conducted a study focusing specifically on
@@
 a taxonomy of code inspection techniques~\cite{laitenberger2002survey}. Johnson conducted
-an investigation into code review practices in open source development and their
+an investigation into code review practices in OSS development and their
 effect on choices made by software project managers~\cite{johnson2006collaboration}.
@@
 More recently, Rigby has done extensive work examining code review practices in
-open source software development~\cite{rigby2012open}.  For example in
-a study of practices in the Apache project~\cite{rigby2008open}
+OSS development~\cite{rigby2008open, Rigb2011a, rigby2012open}.  For example in
+a study of practices in the Apache project~\cite{rigby2008open},
 they data-mined the email archives and found that reviews were typically small
@@
 value to engineers later trying to understand or modify the discussed code.
-They found that ``the meat of the code review dialog, no matter what medium, is
-the articulation of design rationale'' and, thus, ``code reviews are an enticing
-opportunity for capturing design rationale.''
+They found that \quotation{the meat of the code review dialog, no matter what medium, is
+the articulation of design rationale} and, thus, \quotation{code reviews are an enticing
+opportunity for capturing design rationale.}
 

### 2013-03-16 paper.tex
moved sections into a separate dir
done up to section IV (partially included)
working on section 4
done with section 5
did another full pass on the paper
changed according to camera-ready formatting instructions

@@
 \usepackage{float} % defines newfloat used below
-\usepackage[sort&compress]{natbib} 
+%\usepackage[sort&compress]{natbib} % <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< commented out
 %\usepackage{todos} % Use \todoset{all=off} to turn off todos in a paper.
@@
 \usepackage{microtype}
+\usepackage{paralist}
 
@@
 % doing so eliminates compiler warnings.
-\usepackage[bookmarks, colorlinks, citecolor=green, urlcolor=blue,
-			filecolor=blue, linkcolor=blue]{hyperref}
+\usepackage{hyperref}
 % According to the hyperref readme, algorithm must follow hyperref
@@
 \newcommand{\etal}{\hbox{\emph{et al.}}\xspace}
-\newcommand{\eg}{\hbox{\emph{e.g.}}\xspace}
-\newcommand{\ie}{\hbox{\emph{i.e.}}\xspace}
+\newcommand{\eg}{\hbox{\emph{e.g.},}\xspace}
+\newcommand{\ie}{\hbox{\emph{i.e.},}\xspace}
 \newcommand{\scil}{\hbox{\emph{sc.}}\xspace} %scilicet: it is permitted to know
@@
 \newcommand{\vs}{\hbox{\emph{vs.}}\xspace}
+
+\newcommand{\cat}[1]{{\textsc{#1}}}
 
@@
 
-\usepackage{mdframed}
+%\usepackage{mdframed} <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< commented out
 %% Document-specific hyperref options
@@
 
+\newcommand{\figref}[1]{Figure~\ref{#1}\xspace}
+
 \newcounter{copyrightbox}
@@
 		University of Lugano, Switzerland & & Redmond, Washington, USA \\
-		\texttt{alberto.bacchelli@usi.ch} &  &\texttt{cbird@microsoft.com}
+		alberto.bacchelli@usi.ch &  & cbird@microsoft.com
 	\end{tabular}
@@
 
-\input{introduction}
-\input{related_work}
-\input{methodology}
-\input{expectations}
-\input{outcomes}
-\input{challenges}
-\input{implications}
-\input{limitations}
-\input{conclusion}
+\input{sections/introduction}
+\input{sections/related_work}
+\input{sections/methodology}
+\input{sections/expectations}
+\input{sections/outcomes}
+\input{sections/challenges}
+\input{sections/implications}
+\input{sections/limitations}
+\input{sections/conclusion}
 
 \bibliographystyle{abbrv}
+%\bibliographystyle{plain}
 %\setlength{\bibsep}{3pt}

### 2013-03-16 sections/outcomes.tex
Added figures and references to them
changed according to camera-ready formatting instructions

@@
 
+\begin{figure}[t] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=\columnwidth]{comments.pdf}
+   \vspace{-1.5em}
+   \caption{Proportion of comments by card sort category.}
+   \label{fig:comments}
+   \vspace{-1.5em}
+\end{figure}
+
 Our second research question seeks to understand what the actual outcomes of code reviews are, and whether they match the motivations and expectations outlined in the previous section.
 To that end, we conducted indirect field research~\cite{lethbridge2005studying} by analyzing the content of 200 threads (corresponding to 570 comments) recorded by CodeFlow.
-\figref{XXX4} shows the categories of comments found through the card sort.
+\figref{fig:comments} shows the categories of comments found through the card sort.
 
@@
 
-\subsection{Finding defects: When expectations do not meet reality}
+\subsection{Finding Defects: When Expectations Do Not Meet Reality}
 
